---
title: "Import_2023_03_02"
author: Mariia Grybenko
date: today
format:
  html:
    toc: true
    toc-depth: 6
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: false
execute:
  warning: false
  message: false
---

## About

This is the document for importing data file on neuron stimulation from 2023.03.02. 

```{r}
#| include: false
# install.packages("readABF")
library(readABF)
library(tidyverse)
```

## Introducing example file, making tidy, creating functions

```{r - example import}
 

example_file <- (readABF("data/2023-03-02/2023_03_02_0007.abf"))

extract_sweep <- function(sweep_n, abf) {
  abf %>% 
    as.data.frame(sweep = sweep_n) %>%
    mutate(sweep = sweep_n)
}

#extract_sweep(sweep = 1, abf = example_file)

all_sweeps_example <- map(1:length(example_file$data), extract_sweep, abf = example_file)


all_rows_example <- bind_rows(all_sweeps_example)


abf_to_df <- function(abf){
  map(1:length(abf$data), extract_sweep, abf = abf) |> 
    bind_rows()
}

#abf_to_df(example_file)
```

## Importing my data
Transforming into a data frame, changing names, binding data 

```{r - true import}
files <- list.files("data/2023-03-02", pattern = ".abf", full.names = TRUE)
abfs <- map(files, readABF)
dfs <- (map(abfs, abf_to_df))                    


names(dfs) <- files |> str_sub(-9, -5)
# names(dfs) <- str_sub(files, -9, -5) - second option 


dfs <- dfs %>% map(
  rename, time_s = `Time [s]`, current_pA = `Currnt0Pr [pA]`
)

dfs <- map(dfs, select, time_s, current_pA, sweep)

all_sweeps <- bind_rows(dfs, .id = "file" )

          
```

## READ_me on the 2023_03_02 stimulus data 
This is the file with parts of code used to annotate and make tidy 
all imported data from each cell. The code in each chunk was based on
descriptions, stored in the .txt files within each cell data folder

Transforming into a data frame, changing names, selecting variables
```{r}

files <- list.files("data/2023-03-02", pattern = ".abf", full.names = TRUE)

abfs <- map(files, readABF)

# making short names for each object like with last 5 symbols "_0004"
names(abfs) <- files %>% str_sub(-9, -5)

# Transforming abf objects to the tidy dataframes
dfs <- map(abfs, abf_to_df)

# changing names for variables with non-standard symbols
dfs <- map(dfs, rename, time_s = `Time [s]`, current_pA = `Currnt0Pr [pA]`)

# select only important variables:
dfs <- map(dfs, select, time_s, current_pA, sweep)

```

Binding, filtering based on the condition
```{r _2023_03_02}

# 2023-03-02
all_sweeps <- bind_rows(dfs, .id = "file")

all_sweeps <- all_sweeps %>%
    # "_0006" file have changing stimulus and not suitable for compare 
    filter (file != "_0006") %>%
    mutate(
        # type of fibers stimulated in this file / sweep (A, C or both)
        stim = if_else(
            file %in% c("_0005", "_0008", "_0011", "_0013"),
            "A",
            if_else(
                sweep <= 7, "AC", "C"
            )
        ),
        # condition of recording (control, noradrenaline, washout)
        cond = case_when(
            file %in% c("_0004", "_0005", "_0007", "_0008") ~ "control",
            file %in% c("_0010", "_0011") ~ "noradrenalin",
            file %in% c("_0012", "_0013") ~ "washout"
        )
    ) %>%
    mutate(
        stim = as.factor(stim),
        cond = as.factor(cond),
    )

```

## Session info

```{r}
sessionInfo()
```



