---
title: "Import_2023_03_02"
author: Mariia Grybenko
date: today
format:
  html:
    toc: true
    toc-depth: 6
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: false
execute:
  warning: false
  message: false
---
```{r}
```


```{r}
```


```{r}
```


```{r}
```

## About

This is the document for importing data file on neuron stimulation from 2023.03.02. 

```{r}
#| include: false
# install.packages("readABF")
library(readABF)
library(tidyverse)
library(GGally)
```

##  Working with an example file
Making tidy data, creating functions for data frame transformations

```{r - example import}
 

example_file <- (readABF("data/2023-03-02/2023_03_02_0007.abf"))

extract_sweep <- function(sweep_n, abf) {
  abf %>% 
    as.data.frame(sweep = sweep_n) %>%
    mutate(sweep = sweep_n)
}

#extract_sweep(sweep = 1, abf = example_file)

all_sweeps_example <- map(1:length(example_file$data), extract_sweep, abf = example_file)


all_rows_example <- bind_rows(all_sweeps_example)


abf_to_df <- function(abf){
  map(1:length(abf$data), extract_sweep, abf = abf) |> 
    bind_rows()
}

#abf_to_df(example_file)
```

## Importing my data
Transforming into a data frame, changing names, binding data 

```{r - true import}
files <- list.files("data/2023-03-02", pattern = ".abf", full.names = TRUE)
abfs <- map(files, readABF)
dfs <- (map(abfs, abf_to_df))                    


names(dfs) <- files |> str_sub(-9, -5)
# names(dfs) <- str_sub(files, -9, -5) - second option 


dfs <- dfs %>% map(
  rename, time_s = `Time [s]`, current_pA = `Currnt0Pr [pA]`
)

dfs <- map(dfs, select, time_s, current_pA, sweep)

all_sweeps <- bind_rows(dfs, .id = "file" )

          
```

## READ_me on the 2023_03_02 stimulus data 
This is the file with parts of code used to annotate and make tidy 
all imported data from each cell. The code in each chunk was based on
descriptions, stored in the .txt files within each cell data folder

Transforming into a data frame, changing names, selecting variables
```{r}

files <- list.files("data/2023-03-02", pattern = ".abf", full.names = TRUE)

abfs <- map(files, readABF)

# making short names for each object like with last 5 symbols "_0004"
names(abfs) <- files %>% str_sub(-9, -5)

# Transforming abf objects to the tidy dataframes
dfs <- map(abfs, abf_to_df)

# changing names for variables with non-standard symbols
dfs <- map(dfs, rename, time_s = `Time [s]`, current_pA = `Currnt0Pr [pA]`)

# select only important variables:
dfs <- map(dfs, select, time_s, current_pA, sweep)

```

Binding, filtering based on the condition
```{r _2023_03_02}

# 2023-03-02
all_sweeps <- bind_rows(dfs, .id = "file")

all_sweeps <- all_sweeps %>%
    # "_0006" file have changing stimulus and not suitable for compare 
    filter (file != "_0006") %>%
    mutate(
        # type of fibers stimulated in this file / sweep (A, C or both)
        stim = if_else(
            file %in% c("_0005", "_0008", "_0011", "_0013"),
            "A",
            if_else(
                sweep <= 7, "AC", "C"
            )
        ),
        # condition of recording (control, noradrenaline, washout)
        cond = case_when(
            file %in% c("_0004", "_0005", "_0007", "_0008") ~ "control",
            file %in% c("_0010", "_0011") ~ "noradrenalin",
            file %in% c("_0012", "_0013") ~ "washout"
        )
    ) %>%
    mutate(
        stim = as.factor(stim),
        cond = as.factor(cond),
    )

```

## Session info

```{r}
sessionInfo()
```


## Calculating passive parameters

Calculated the next passive parameters: baseline [pA], peak [pA], plateau [pA], membrane resistance [MOmh], series resistance [MOhm], membrane capacity [pF]. Created a joined table with all passive parameters. Created an interactive plot with ggplotly. Created a visual (correlations/clusters) with GGally. 

```{r}
#omitting random peaks
voltage_step_data <- all_sweeps |>
  filter(between(time_s, 0.12, 0.5))

#setting time frames 
baseline_tw <- c(0.14, 0.16)
peak_tw <- c(0.159, 0.165)
plateau_tw <- c(0.175, 0.2)


#plotting data on the sweeps 
plot_voltage_step <- ggplot(voltage_step_data, 
       aes(x = time_s, y = current_pA, group = interaction(file, sweep), color = file),
       )+
  geom_vline(xintercept = baseline_tw, color = "lightblue") +
  geom_vline(xintercept = peak_tw, color = "pink") +
  geom_vline(xintercept = plateau_tw, color = "lightgreen") +
  geom_line(alpha = 0.6)
  
#adding plotly, making the plot interactive
ggplotly(plot_voltage_step)


passive_parameters <- voltage_step_data |>
  group_by(file,sweep) |>
  summarise(
    baseline_pA = mean(current_pA[between(time_s, baseline_tw[1], baseline_tw[2])]),
    peak_pA = min(current_pA[between(time_s, peak_tw[1], peak_tw[2])]) - baseline_pA,
    plateau_pA = mean(
      current_pA[between(time_s, plateau_tw[1], plateau_tw[2])]
    ) - baseline_pA,
    Rs_MOhm = -10000 / peak_pA,
    Rtotal_MOhm = -10000 / plateau_pA,
    Rmemb_MOhm = Rtotal_MOhm - Rs_MOhm
    )


#### Calculating slope, the tau from the slope and consequently the membrane capacity 

decay_tw <- c(0.16, 0.175)

decay_fits <- voltage_step_data |>
  filter(time_s >= decay_tw[1], time_s <= decay_tw[2]) |>
  nest(.by = c(file, sweep)) |>
# now we have list-column "data" which will be used for modelling
  mutate(
    models = map(
      data,
      possibly(\(x) nls(x,
                        formula = current_pA ~ SSasymp(time_s, Asym, R0, lrc))
               )
      ),
# using SSasymp() as in this tutorial:
# https://douglas-watson.github.io/post/2018-09_exponential_curve_fitting/
# but wrapping with possible() to still run if exponent does not fit
coefs = map(models, coef), #extracting coeficients from model
lrc = map_dbl(coefs, \(x) pluck(x, "lrc", .default = NA)),
# pluck() extract value of the lrc from all coefs lists.
tau = 1/exp(lrc)
# SSasymp calculate natural Logarithm of Rate Constant (lrc),
# but we need tau which is 1 / rate constant
) |>
  select(file, sweep, tau) # only the variable we needed

#### END OF HARD PART ####

#adding membrane capacity to passive parameters 
passive_parameters <- passive_parameters |> 
  left_join(decay_fits)|>
  mutate(
    Cmemb_approx_pF = tau / Rs_MOhm * 10^6,
    Cmemb_pF = tau / (1/(1/Rs_MOhm +1/Rmemb_MOhm)) * 10^6
    
  )

#creating the table with selected passive parameters
passive_parameters |>
  select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
  datatable(rownames = FALSE)

#creating visuals instead of a busy data table
passive_parameters |>
  select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
  na.omit() |>
  GGally::ggpairs(
    aes(color = file),
    columns = 3:6
    )
           

    
```

```


## Mono- and Polysynaptic outputs 
Моносинаптичні відповіді 
- швидка відповідь пісял стимуляції (менше 15-20 мс)
- синхронні (менше 1 мс)
- на кожному свіпу, низький failure rate 

Вирізняємо моносинаптичні, аби розуміти де діє норадреналін, вказує на механізм. 

## Postsynaptic responces to AC stimulation 

```
```{r}
AC_responces <- all_sweeps |>
  filter(between(time_s, 0.28, 0.4)) |>
  filter(stim == "AC")

bl_area_tw <- c(0.2934, 0.2964)
area_tw <- c(0.316, 0.4)

ACtraces_plot <- ggplot(
    AC_responces,
    aes(
        x = time_s,
        y = current_pA,
        group = interaction(file, sweep),
        color = file
    )) +
    geom_line(
        alpha = 0.4,
        show.legend = FALSE
        ) +
    geom_vline(xintercept = bl_area_tw, color = "lightblue") +
    geom_vline(xintercept = area_tw, color = "pink")

ggplotly(ACtraces_plot)

AC_areas <- AC_responces |>
  group_by(file, sweep) |>
  mutate(
    current_pA = current_pA - mean(current_pA[between(time_s, bl_area_tw[1], bl_area_tw[2])])
  ) |>
  filter(between(time_s, area_tw[1], area_tw[2])) |>
  summarize(
  total_area_pAs =  sum(current_pA) * 0.00005
)

```


```{r}

calculate_response <- function(data, bl_tw, resp_tw, param) {
  data |>
    group_by(file, sweep) |>
    mutate(
      current_pA = current_pA - 
        mean(current_pA[between(time_s, bl_tw[1], bl_tw[2])])
    ) |>
    filter(between(time_s, resp_tw[1], resp_tw[2])) |>
    summarise(
      parameter = case_when(
        param == "amplitude" ~ min(current_pA),
        param == "area" ~ sum(current_pA) * 0.0005,
        .unmatched = "error"
        
        
      )
    )
}

```


## Performing the monosynapses calculations (not found)
```{r}
#NO real monosynapses, performed the calculations simply for practice 


# bl_mono2_tw <- c(0.3235, 0.3236)
# mono2_tw <- c(0.32520, 0.32550)
# 
# AC_mono_1_ampl <- calculate_response(
#   AC_responces,
#   bl_tw = bl_mono2_tw,
#   resp_tw = mono2_tw,
#   param = "amplitude"
# )
# 
# 
# 
# bl_mono1_tw <- c(0.3230, 0.32305)
# mono1_tw <- c(0.33435, 0.33550)
# 
# AC_mono_2_ampl <- calculate_response(
#   AC_responces,
#   bl_tw = bl_mono1_tw,
#   resp_tw = mono1_tw,
#   param = "amplitude"
# )


AC_areas <- AC_areas |>
  rename(area_pAs = total_area_pAs)

AC_mono_1_ampl <- AC_mono_1_ampl |>
  rename(mono1_amp_pA = parameter)
AC_mono_2_ampl <- AC_mono_2_ampl |>
  rename(mono2_amp_pA = parameter)


# Joining the 3 tibbles into the main data window
# final_data_test <- window_df |>
#   left_join(AC_mono_1_ampl, by = c("file", "sweep")) |>
#   left_join(AC_mono_2_ampl, by = c("file", "sweep")) |>
#   left_join(AC_areas, by = c("file", "sweep"))


AC_all_parameters <- window_df |>
  left_join(AC_areas, by = c("file", "sweep")) |>
  left_join(passive_parameters, by = c("file", "sweep"))

write_tsv(AC_all_parameters, file = "../AC_all_parameters.tsv")
```

```


