---
title: "Importing_data"
author: "maryna"
date: July 20, 2026
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: true
execute:
  warning: false
  message: false
---

## Importing data from cell 2023_03_22_2

This is .qmd document for importing patch clamp data for cell 2023_03_22_2

Cell: 2023_03_22_2

*(Shortcut for new chunk: ctrl+alt+I)*

The chunk below was made to make `dplyr` overwrite system's `rename` and make `as.factor` work as base one.

```{r setup}
library(readABF)
library(conflicted)
library(dplyr)
library(tidyverse)

# Tell R to prefer dplyr functions over base/stats
conflict_prefer("rename", "dplyr")
conflict_prefer("filter", "dplyr")
conflicts_prefer(base::as.factor)

```

Firstly, import one file. To make data tidy we use `as.data.frame.`The file contains several sweeps, so we need to specify which one. Here `sqrt` is an argument and not a function, therefore no brackets needed. Function `map` uses the second argument as a function that applies to the first one, `map` always return a list.

```{r import_file}
example_file <- readABF("../data/2023-03-22_2/2023_03_22_0032.abf")

#map(1:10, sqrt)

extract_sweep <- function(sweep_n, abf){
  abf |> 
    as.data.frame(sweep = sweep_n) |> 
    mutate(sweep = sweep_n) 
# We add a variable to the given data frame, which will be usefull
}

#extract_sweep(5, example_file)
sweep_dfs <- map(1:length(example_file$data), 
                 extract_sweep, abf = example_file)
example_df <- bind_rows(sweep_dfs)
# To combine data frames created by map to one

abf_to_df <- function(abf_given){
  map(1:length(abf_given$data), extract_sweep, abf = abf_given) |> 
    bind_rows()
    
}
#abf_to_df(example_file)
knitr::kable(head(example_df))
```

Import all data for the cell.

Function list.files with parameter `full.names = TRUE` it saves the pathes and not just names. `pattern` filters the files by types

```{r import}
files <- list.files("../data/2023-03-22_2/", 
           full.names = TRUE, 
           pattern = ".abf")

abfs <- map(files, readABF)

dfs <- map(abfs, abf_to_df)

files |> str_sub(-9, -5) 
# We get the files' endings to use it as names

names(dfs) <- files |> 
  str_sub(-9, -5)
#names(dfs)

# To make the previous name be seen by R we take it in ``, because 
# it was written badly (with space and special symbols) 
example_df <- example_df |> rename(
  time_s = `Time [s]`,
  current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]" ))
)

dfs <- map(
  dfs, 
  rename, 
  time_s = `Time [s]`,
  current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]" ))
  )
dfs <- map(dfs, select, time_s, current_pA, sweep)

all_sweeps <- bind_rows(dfs, .id = "file")
all_sweeps <- all_sweeps |> 
  # only the files with AC or C stim, there is no recordings of 
  # A-stim in Noradrenaline so no need to import it
  filter (file %in% c("_0032", "_0036", "_0037")) |> 
  mutate(
    # type of fibers stimulated in this file / sweep (C or both)
    stim = if_else(sweep <= 10, "AC", "C"),
    # condition of recording (control, noradrenaline, washout)
    cond = case_when(
      file == "_0032" ~ "control",
      file == "_0036" ~ "noradrenalin",
      file == "_0037" ~ "washout"
    )
  ) |> 
  mutate(
    stim = as.factor(stim),
    cond = as.factor(cond),
  )
```

## Session info

```{r info}
sessionInfo()
```
