---
title: "Lamin X project"
author: "Maryna D."
date: "20-23, July 2026"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: false
execute:
  warning: false
  message: false
---

## Patch clamp data import

Cell: `2023_03_22_2`

This Quarto (.qmd) document outlines the workflow for importing, tidying, and analyzing raw electrophysiological patch-clamp recordings from cell 2023_03_22_2.

```{r setup}
library(readABF)
library(conflicted)
library(dplyr)
library(DT)
library(tidyverse)
library(plotly)

# Set conflict resolutions to prefer dplyr methods over base/stats
conflict_prefer("rename", "dplyr")
conflict_prefer("filter", "dplyr")
conflicts_prefer(base::as.factor)
#-

```

```{r import_file}

# Import one example file
example_file <- readABF("../data/2023-03-22_2/2023_03_22_0032.abf")

# make data tidy
# as.data.frame(example_file, sweep = 1)

extract_sweep <- function(sweep_n, abf){
    abf |>
        as.data.frame(sweep = sweep_n) |>
        mutate(sweep = sweep_n)
}

# extract_sweep(sweep_n = 1, abf = example_file)

sweep_dfs <- map(1:length(example_file$data), extract_sweep, abf = example_file)

example_df <- bind_rows(sweep_dfs)

abf_to_df <- function(abf){
    map(1:length(abf$data), extract_sweep, abf = abf) |>
        bind_rows()
}

example_df <- abf_to_df(example_file)

datatable(head(example_df, 1000))

knitr::kable(head(example_df))
```

## Now import all data for the cell

```{r real_import}

files <- list.files(
    "../data/2023-03-22_2/", 
    pattern = ".abf", 
    full.names = TRUE
    )

# extract the 4-digit file id, e.g. ".../2023_03_22_0032.abf" -> "0032"
# (regex is more robust than str_sub here since we don't know if the
# filenames end in "_0032.abf" or "0032_.abf" etc.)
file_ids <- str_extract(basename(files), "\\d{4}(?=_?\\.abf$)")

abfs <- map(files, readABF)

dfs <- map(abfs, abf_to_df)

names(dfs) <- file_ids

# the exact channel names readABF produces depend on the signals recorded
# for this cell (Currnt0Pr, Volt2Pr per the ATF header) -- any_of() means
# this still works if the naming differs slightly between files
dfs <- map(
    dfs, 
    rename, time_s = `Time [s]`, 
    current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]", "Volt2Pr [pA]"))
    )

dfs <- map(dfs, select, time_s, current_pA, sweep)

all_sweeps <- bind_rows(dfs, .id = "file")

datatable(head(all_sweeps, 10000))

```

## Annotate

Here we add the variables for stimulation type and condition (control, noradrenaline, washout), per the notebook entry above.

```{r annotate}

all_sweeps <- all_sweeps %>%
    filter(file %in% c("0032", "0036", "0037")) %>%
    mutate(
        # type of fibers stimulated in this file / sweep (C or both)
        stim = if_else(sweep <= 10, "AC", "C"),
        # condition of recording (control, noradrenaline, washout)
        cond = case_when(
            file == "0032" ~ "control",
            file == "0036" ~ "noradrenalin",
            file == "0037" ~ "washout"
        )
    ) %>%
    mutate(
        stim = as.factor(stim),
        cond = as.factor(cond),
    )

datatable(head(all_sweeps, 100))

```

```{r saving data}

write_tsv(all_sweeps, file = "../all_sweeps.tsv")

# saving only part of the data for shiny app to make it faster
write_tsv(all_sweeps |> filter(between(time_s, 0.1, 0.7)), 
file = "../all_sweeps_part.tsv")

```
## Calculate passive parameters

We calculated the next passive parameters :

baseline (pA), serial resistance

```{r passive}

# saving only part of data with voltage step to work with
voltage_step_data <- all_sweeps |>
    filter(between(time_s, 0.12, 0.22))

#setting time windows
baseline_tw <- c(0.14, 0.16)
peak_tw <- c(0.159, 0.165)
plateau_tw <- c(0.180, 0.198)
decay_tw <- c(0.1605, 0.175)

# checking time windows
plot_voltage_step <- ggplot(voltage_step_data, 
       aes(x = time_s, y = current_pA, group = interaction(file,sweep), color = file)
       )+
    geom_line(alpha = 0.6) +
    geom_vline(xintercept = baseline_tw, color = "lightblue") +
    geom_vline(xintercept = peak_tw, color = "pink") +
    geom_vline(xintercept = plateau_tw, color = "lightgreen")+
    geom_vline(xintercept = decay_tw, color = "gold")

ggplotly(plot_voltage_step)

# calculating resistances
passive_parameters <- voltage_step_data |>
    group_by(file,sweep) |>
    summarise(
        baseline_pA = mean(
            current_pA[between(time_s, baseline_tw[1], baseline_tw[2])]
            ),
        peak_pA = min(
            current_pA[between(time_s, peak_tw[1], peak_tw[2])]
            ) - baseline_pA,
        plateau_pA = mean(
            current_pA[between(time_s, plateau_tw[1], plateau_tw[2])]
            ) - baseline_pA,
        Rs_MOhm = -10000 / peak_pA,
        Rtotal_MOhm = -10000 / plateau_pA,
        Rmemb_MOhm = Rtotal_MOhm - Rs_MOhm
        )


decay_tw <- c(0.1605, 0.175)

#### This Part of code will be VERY hard to explain ####
# so we will just copypaste it (it must be robust to errors)
# and i will give a general overview what it is doing:

decay_fits <- voltage_step_data |> 
    filter(time_s >= decay_tw[1], time_s <= decay_tw[2]) |> 
    nest(.by = c(file, sweep)) |>
    # now we have list-column "data" which will be used for modelling
    mutate(
        models = map(
            data, 
            possibly(\(x) nls(x, 
                              formula = current_pA ~ SSasymp(time_s, Asym, R0, lrc))
                     )
            ),
        # using SSasymp() as in this tutorial:
        # https://douglas-watson.github.io/post/2018-09_exponential_curve_fitting/
        # but wrapping with possible() to still run if exponent does not fit
        coefs = map(models, coef), #extracting coeficients from model
        lrc = map_dbl(coefs, \(x) pluck(x, "lrc", .default = NA)), 
        # pluck() extract value of the lrc from all coefs lists.
        tau = 1/exp(lrc) 
        # SSasymp calculate natural Logarithm of Rate Constant (lrc),
        # but we need tau which is 1 / rate constant
        ) |>
    select(file, sweep, tau) # only the variable we needed

#### END OF HARD PART ####

passive_parameters <- passive_parameters |>
    left_join(decay_fits) |>
    mutate(
        Cmemb_approx_pF = tau / Rs_MOhm * 10^6,
        # tau (s) / R (MOhms) = C (uF), *10^6 to convert to pF
        
        Cmemb_pF = tau / (1/(1/Rs_MOhm + 1/Rmemb_MOhm)) * 10^6
        # actually tau depends on both Rseries and Rmemb, 
        # and from Capacitor point of view they are in parallel 
        )


datatable(passive_parameters)


# show calculated passive parameters
passive_parameters |>
    select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    datatable(rownames = FALSE)

# and nice plot to see them
passive_parameters |>
      select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    na.omit() |>
    GGally::ggpairs(
          aes(color = file), 
          columns = 3:6
          )
```

## Calculating the postsynaptic responces

### postsynaptic responces to AC stimulation

```{r ACpostsynaptic}

# saving only part of data with voltage step to work with
AC_responses <- all_sweeps |>
    filter(between(time_s, 0.28, 0.4)) |>
    filter(stim == "AC")

# setting time windows for total area of response
bl_area_tw <- c(0.285, 0.2935)
area_tw <- c(0.3045, 0.384)

# quick plotly to check time windows
ACtraces_plot <- ggplot(AC_responses, aes(
    x = time_s, 
    y = current_pA,
    group = interaction(file, sweep),
    color = file
    ))+
    geom_line(
        alpha = 0.5,
        show.legend = FALSE
        ) +
    geom_vline(xintercept = bl_area_tw, color = "lightblue") +
    geom_vline(xintercept = area_tw, color = "gold")

ggplotly(ACtraces_plot)


calculate_response <- function(data, bl_tw, resp_tw, param) {
    data |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA -
            mean(current_pA[between(time_s, bl_tw[1], bl_tw[2])])
    ) |>
    filter(between(time_s, resp_tw[1], resp_tw[2])) |>
    summarise(
        parameter = case_when(
            param == "amplitude" ~ min(current_pA),
            param == "area" ~ sum(current_pA) * 0.00005,
            .default = NA_real_
            )
    )
}

bl_mono1_tw <- c(0.3035, 0.3045)
mono1_tw <- c(0.3075, 0.3085)

AC_mono1_aplitudes <- calculate_response(
    AC_responses,
    bl_tw = bl_mono1_tw,
    resp_tw = mono1_tw,
    param = "amplitude"
    )

AC_areas <- calculate_response(
    AC_responses,
    bl_tw = bl_area_tw,
    resp_tw = area_tw,
    param = "area"
    )

AC_areas <- AC_areas |>
    rename(area_pAs = parameter)

AC_mono1_aplitudes <- AC_mono1_aplitudes |>
    rename(mono1_amp_pA = parameter)

AC_all_params <- AC_areas |> 
    left_join(AC_mono1_aplitudes) |>
    left_join(passive_parameters)

write_tsv(AC_all_params, file = "../AC_all_params.tsv")

```

### postsynaptic responces to C stimulation

Sweeps 11+ in every file are C-fiber-only stimulation.

The AC-derived placeholder windows (0.314-0.318 s) do NOT work for the C-fiber
data -- checking the raw traces shows that window only captures the tail of the
stimulus artifact (a fast, essentially jitter-free deflection that peaks at
exactly 0.3111 s in *every single sweep*, which is the signature of an
electrical artifact rather than a synaptic event). The genuine C-fiber
response is slower and, consistent with the slower and more variable
conduction of unmyelinated C-fibers, its peak latency jitters sweep-to-sweep
(peak times ranging roughly 0.32-0.365 s across sweeps/files, vs. the
artifact's fixed 0.3111 s). The windows below are based on inspecting the raw
sweeps directly:

- baseline: `c(0.295, 0.308)` -- flat period just before the stimulus artifact
- response window: `c(0.32, 0.5)` -- starts after the artifact has decayed and
  covers the full rise/peak/decay of the C-fiber response (much longer-lasting
  than the AC monosynaptic component, so the window is wider than the AC one)

Because the peak latency jitters from sweep to sweep, we use one window (not a
tight ~1 ms one like AC's `mono1_tw`) and take the minimum within it, the same
way `calculate_response()` already works for `param = "amplitude"`. Check
`Ctraces_plot` below against these vlines before trusting the numbers -- adjust
if your data looks different.

```{r Cpostsynaptic}

# saving only part of data to work with -- wide enough to show baseline,
# the stimulus artifact, and the full C-fiber response decay
C_responses <- all_sweeps |>
    filter(between(time_s, 0.29, 0.55)) |>
    filter(stim == "C")

# setting time windows (see note above for how these were chosen)
C_bl_tw   <- c(0.295, 0.308)
C_resp_tw <- c(0.32, 0.5)

# quick plotly to check time windows
Ctraces_plot <- ggplot(C_responses, aes(
    x = time_s,
    y = current_pA,
    group = interaction(file, sweep),
    color = file
    ))+
    geom_line(
        alpha = 0.5,
        show.legend = FALSE
        ) +
    geom_vline(xintercept = C_bl_tw, color = "lightblue") +
    geom_vline(xintercept = C_resp_tw, color = "gold")

ggplotly(Ctraces_plot)

# total area of the C-fiber response
C_areas <- calculate_response(
    C_responses,
    bl_tw = C_bl_tw,
    resp_tw = C_resp_tw,
    param = "area"
    ) |>
    rename(area_pAs = parameter)

# peak amplitude of the C-fiber response -- a single wide window (not a tight
# ~1 ms one like AC's mono1_tw), since the peak latency jitters sweep to sweep
C_peak_amplitudes <- calculate_response(
    C_responses,
    bl_tw = C_bl_tw,
    resp_tw = C_resp_tw,
    param = "amplitude"
    ) |>
    rename(peak_amp_pA = parameter)

C_all_params <- C_areas |>
    left_join(C_peak_amplitudes, by = c("file", "sweep")) |>
    left_join(passive_parameters, by = c("file", "sweep"))

write_tsv(C_all_params, file = "../C_all_params.tsv")

datatable(C_all_params)

```

## Filtering our data from parameters

```{r filtering}
# selecting only the most important variables
passive_parameters <- passive_parameters |>
      select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF)

# setting limits
baseline_lims <- c(-200,0)
Rs_lims <- c(
    0.9 * median(passive_parameters$Rs_MOhm, na.rm = TRUE),
    1.1 * median(passive_parameters$Rs_MOhm, na.rm = TRUE)
    )
Rm_lims <- c(
    0.5 * median(passive_parameters$Rmemb_MOhm, na.rm = TRUE),
    2 * median(passive_parameters$Rmemb_MOhm, na.rm = TRUE)
    )
Cm_lims <- c(
    0.5 * median(passive_parameters$Cmemb_pF, na.rm = TRUE),
    2 * median(passive_parameters$Cmemb_pF, na.rm = TRUE)
    )

# do quality check -- both per-parameter flags (for plotting) and
# the combined flag (for actually keeping/dropping sweeps)
passive_parameters <- passive_parameters |>
    ungroup() |>
    mutate(
        QC_baseline = between(baseline_pA, baseline_lims[1], baseline_lims[2]),
        QC_Rs       = between(Rs_MOhm, Rs_lims[1], Rs_lims[2]),
        QC_Rm       = between(Rmemb_MOhm, Rm_lims[1], Rm_lims[2]),
        QC_Cm       = between(Cmemb_pF, Cm_lims[1], Cm_lims[2]),
        QC = QC_baseline & QC_Rs & QC_Rm & QC_Cm
    )

# visualisation of quality check -- colored by the criterion specific
# to each facet, so a point's color always matches its own threshold lines
plot_filtering <- ggplot(
    data = passive_parameters |>
        mutate(
            bl_pc = baseline_pA / median(baseline_pA, na.rm = TRUE),
            rs_pc = Rs_MOhm / median(Rs_MOhm, na.rm = TRUE),
            rm_pc = Rmemb_MOhm / median(Rmemb_MOhm, na.rm = TRUE),
            cm_pc = Cmemb_pF / median(Cmemb_pF, na.rm = TRUE)
        ) |>
        select(file, sweep, bl_pc, rs_pc, rm_pc, cm_pc,
               QC_baseline, QC_Rs, QC_Rm, QC_Cm) |>
        pivot_longer(
            cols = c(bl_pc, rs_pc, rm_pc, cm_pc),
            names_to = "param",
            values_to = "percentage"
        ) |>
        mutate(
            QC_param = case_when(
                param == "bl_pc" ~ QC_baseline,
                param == "rs_pc" ~ QC_Rs,
                param == "rm_pc" ~ QC_Rm,
                param == "cm_pc" ~ QC_Cm
            )
        ),
    aes(x = param, y = percentage)
    ) +
    geom_boxplot(outliers = FALSE) +
    geom_point(
        aes(group = interaction(file, sweep), color = QC_param),
        position = position_dodge(0.2)
        ) +
    geom_line(
        aes(group = interaction(file, sweep)),
        position = position_dodge(0.2),
        alpha = 0.2
        ) +
    annotate(
        "segment",
        x = c(0.6, 1.6, 1.6, 2.6, 2.6, 3.6, 3.6),
        xend = c(1.4, 2.4, 2.4, 3.4, 3.4, 4.4, 4.4),
        y = c(-200/median(passive_parameters$baseline_pA), 0.5, 2, 0.5, 2, 0.9, 1.1),
        yend = c(-200/median(passive_parameters$baseline_pA), 0.5, 2, 0.5, 2, 0.9, 1.1),
        color = "darkblue"
        ) +
    ylim(0, 4) +
    labs(color = "QC")

ggplotly(plot_filtering)

# filtering raw sweeps data -- uses the COMBINED QC (all 4 criteria),
# not the per-parameter one used just for the plot's coloring
keep <- passive_parameters %>% filter(QC) %>% select(file, sweep)
sweeps_keep <- all_sweeps |>
    inner_join(keep, by = c("file", "sweep"))

```

## Total area of response by condition

```{r AC_condition_avg_plot}

# baseline-subtracted traces for every sweep, for plotting
AC_area_plot <- AC_responses |>
    filter(between(time_s, 0.28, 0.4)) |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA -
            mean(current_pA[between(time_s, bl_area_tw[1], bl_area_tw[2])])
    ) |>
    ungroup()

# condition-level average trace (mean across sweeps at each timepoint)
AC_average_plot <- AC_area_plot |>
    group_by(cond, time_s) |>
    summarise(current_pA = mean(current_pA), .groups = "drop")

ACtraces_plot <- ggplot(
    AC_area_plot,
    aes(
        x = time_s,
        y = current_pA,
        color = cond
        )
    ) +
    geom_line(
        aes(group = interaction(file, sweep)),
        alpha = 0.1,
        show.legend = FALSE
        ) +
    geom_line(
        data = AC_average_plot,
        aes(group = cond)
        ) +
    geom_vline(xintercept = bl_area_tw, color = "lightblue") +
    geom_vline(xintercept = area_tw, color = "gold") +
    ylim (-300, 20) +
    labs(x = "Time (s)", y = "Current (pA)",
         color = "Condition")

ggplotly(ACtraces_plot)

```

### One-Way ANOVA for Area and Amplitude

```{r statistical_analysis}
library(stats)
# Fit ANOVA models for response area and first monosynaptic peak amplitude
# Re-create AC_all_params with the 'cond' variable restored
AC_all_params <- AC_areas |> 
    left_join(AC_mono1_aplitudes, by = c("file", "sweep")) |>
    left_join(passive_parameters, by = c("file", "sweep")) |>
    mutate(
        cond = case_when(
            file == "0032" ~ "control",
            file == "0036" ~ "noradrenalin",
            file == "0037" ~ "washout"
        ),
        cond = as.factor(cond)
    )

# Fit ANOVA models for response area and first monosynaptic peak amplitude
anova_area <- aov(area_pAs ~ cond, data = AC_all_params)
anova_amp  <- aov(mono1_amp_pA ~ cond, data = AC_all_params)

# Display ANOVA summary tables
summary(anova_area)
summary(anova_amp)

# Post-hoc analysis (Tukey's Honest Significant Difference)
tukey_area <- TukeyHSD(anova_area)
tukey_amp  <- TukeyHSD(anova_amp)

tukey_area
tukey_amp
# Plot area and monosynaptic amplitude differences across conditions
p_area <- ggplot(AC_all_params, aes(x = cond, y = area_pAs, fill = cond)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.5) +
    geom_jitter(width = 0.15, size = 2, alpha = 0.8) +
    theme_minimal() +
    labs(
        title = "AC Response Area by Condition",
        x = "Condition",
        y = "Area (pA·s)"
    ) +
    theme(legend.position = "none")

p_amp <- ggplot(AC_all_params, aes(x = cond, y = mono1_amp_pA, fill = cond)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.5) +
    geom_jitter(width = 0.15, size = 2, alpha = 0.8) +
    theme_minimal() +
    labs(
        title = "AC Monosynaptic Amplitude by Condition",
        x = "Condition",
        y = "Amplitude (pA)"
    ) +
    theme(legend.position = "none")

plotly::subplot(ggplotly(p_area), ggplotly(p_amp), titleY = TRUE, margin = 0.05)
```

## Session info

```{r info}

sessionInfo()

```
