---
title: "Import_2025_05_02"
author: "Ivan Blashchak"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: true
    
execute:
  warning: false
  message: false
---

## About

This is .Qmd document for importing patch clamp data for cell 2025_05_02


```{r setup}

# install.packages("readABF")
library(readABF)
library(tidyverse)
library(plotly)
library(DT)

```

## Cell: 2025_05_02

First lets just import one file

```{r import_file}

#import one example file
example_file <- readABF("../data/2025_05_02/2025_05_02_0005.abf")

# make data tidy
# as.data.frame(example_file, sweep = 1)

extract_sweep <- function(sweep_n, abf){
    abf |>
        as.data.frame(sweep = sweep_n) |>
        mutate(sweep = sweep_n)
}

# extract_sweep(sweep_n = 1, abf = example_file)

sweep_dfs <- map(1:length(example_file$data), extract_sweep, abf = example_file)

example_df <- bind_rows(sweep_dfs)

abf_to_df <- function(abf){
    map(1:length(abf$data), extract_sweep, abf = abf) |>
        bind_rows()
}

example_df <- abf_to_df(example_file)

datatable(head(example_df, 10000))

knitr::kable(head(example_df))
```

## Now import all data for the cell

``` {r real_import}

files <- list.files(
    "../data/2025_05_02/", 
    pattern = ".abf", 
    full.names = TRUE
    )

abfs <- map(files, readABF)

dfs <- map(abfs, abf_to_df)

files |> str_sub(-9,-5)

names(dfs) <- str_sub(files, -9,-5)

# example_df |> rename(
#     time_s = `Time [s]`,
#     current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
# )

dfs <- map(
    dfs, 
    rename, time_s = `Time [s]`, 
    current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
    )

dfs <- map(dfs, select, time_s, current_pA, sweep)

all_sweeps <- bind_rows(dfs, .id = "file")

datatable(head(all_sweeps, 10000))

```

## Annotate

Here we add the variables for stimulation type and 
condition (control, noradrenaline, washout)

```{r annotate}


all_sweeps <- all_sweeps %>%
    # for "_0006" only recordings before NA addition were included
    filter ( ! (file == "_0006" & sweep > 10)) %>%
    mutate(
        # type of fibers stimulated in this file / sweep (A, C or both)
        # in this cell only AC simulations were performed
        stim = case_when(
            file == "_0005" ~ if_else(sweep <= 17, "AC", "C"),
            file == "_0006" ~ "C",
            file == "_0007" ~ if_else(sweep <= 18, "AC", "C"),
            file == "_0008" ~ "AC"
        ),
        # condition of recording (control, noradrenaline, washout)
        cond = case_when(
            file %in% c("_0005", "_0006") ~ "control",
            file == "_0007" ~ "noradrenalin",
            file == "_0008" ~ "washout"
        )
    ) %>%
    mutate(
        stim = as.factor(stim),
        cond = as.factor(cond),
    )


datatable(head(all_sweeps, 10000))


```


```{r saving data}

write_tsv(all_sweeps, file = "../all_sweeps.tsv")


# saving only part of the data for shiny app to make it faster
write_tsv(all_sweeps |> filter(between(time_s, 0.1, 0.7)), 
file = "../all_sweeps_part.tsv")

```


## Calculate passive parameters

We calculated the next passive parameters :

baseline (pA), serial resistance


```{r passive}

# saving only part of data with voltage step to work with
voltage_step_data <- all_sweeps |>
    filter(between(time_s, 0.12, 0.22))

#setting time windows
baseline_tw <- c(0.14, 0.16)
peak_tw <- c(0.159, 0.165)
plateau_tw <- c(0.180, 0.198)
decay_tw <- c(0.1605, 0.175)

# checking time windows
plot_voltage_step <- ggplot(voltage_step_data, 
       aes(x = time_s, y = current_pA, group = interaction(file,sweep), color = file)
       )+
    geom_line(alpha = 0.6) +
    geom_vline(xintercept = baseline_tw, color = "lightblue") +
    geom_vline(xintercept = peak_tw, color = "pink") +
    geom_vline(xintercept = plateau_tw, color = "lightgreen")+
    geom_vline(xintercept = decay_tw, color = "gold")

ggplotly(plot_voltage_step)


# calculating resistances
passive_parameters <- voltage_step_data |>
    group_by(file,sweep) |>
    summarise(
        baseline_pA = mean(
            current_pA[between(time_s, baseline_tw[1], baseline_tw[2])]
            ),
        peak_pA = min(
            current_pA[between(time_s, peak_tw[1], peak_tw[2])]
            ) - baseline_pA,
        plateau_pA = mean(
            current_pA[between(time_s, plateau_tw[1], plateau_tw[2])]
            ) - baseline_pA,
        Rs_MOhm = -10000 / peak_pA,
        Rtotal_MOhm = -10000 / plateau_pA,
        Rmemb_MOhm = Rtotal_MOhm - Rs_MOhm
        )


decay_tw <- c(0.1605, 0.175)

#### This Part of code will be VERY hard to explain ####
# so we will just copypaste it (it must be robust to errors)
# and i will give a general overview what it is doing:

decay_fits <- voltage_step_data |> 
    filter(time_s >= decay_tw[1], time_s <= decay_tw[2]) |> 
    nest(.by = c(file, sweep)) |>
    # now we have list-column "data" which will be used for modelling
    mutate(
        models = map(
            data, 
            possibly(\(x) nls(x, 
                              formula = current_pA ~ SSasymp(time_s, Asym, R0, lrc))
                     )
            ),
        # using SSasymp() as in this tutorial:
        # https://douglas-watson.github.io/post/2018-09_exponential_curve_fitting/
        # but wrapping with possible() to still run if exponent does not fit
        coefs = map(models, coef), #extracting coeficients from model
        lrc = map_dbl(coefs, \(x) pluck(x, "lrc", .default = NA)), 
        # pluck() extract value of the lrc from all coefs lists.
        tau = 1/exp(lrc) 
        # SSasymp calculate natural Logarithm of Rate Constant (lrc),
        # but we need tau which is 1 / rate constant
        ) |>
    select(file, sweep, tau) # only the variable we needed

#### END OF HARD PART ####

passive_parameters <- passive_parameters |>
    left_join(decay_fits) |>
    mutate(
        Cmemb_approx_pF = tau / Rs_MOhm * 10^6,
        # tau (s) / R (MOhms) = C (uF), *10^6 to convert to pF
        
        Cmemb_pF = tau / (1/(1/Rs_MOhm + 1/Rmemb_MOhm)) * 10^6
        # actually tau depends on both Rseries and Rmemb, 
        # and from Capacitor point of view they are in parallel 
        )


datatable(passive_parameters)


# show calculated passive parameters
passive_parameters |>
    select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    datatable(rownames = FALSE)

# and nice plot to see them
passive_parameters |>
      select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    na.omit() |>
    filter(file != "_0008") |>
    GGally::ggpairs(
          aes(color = file), 
          columns = 3:6
          )
```

## Calculating the postsynaptic responces

### postsynaptic responces to AC stimulation

```{r ACpostsynaptic}

# saving only part of data with voltage step to work with
AC_responses <- all_sweeps |>
    filter(between(time_s, 0.28, 0.4)) |>
    filter(stim == "AC")

# setting time windows for total area of response
bl_area_tw <- c(0.285, 0.2935)
area_tw <- c(0.3045, 0.384)

# quick plotly to check time windows
ACtraces_plot <- ggplot(AC_responses, aes(
    x = time_s, 
    y = current_pA,
    group = interaction(file, sweep),
    color = file
    ))+
    geom_line(
        alpha = 0.5,
        show.legend = FALSE
        ) +
    geom_vline(xintercept = bl_area_tw, color = "lightblue") +
    geom_vline(xintercept = area_tw, color = "gold")

ggplotly(ACtraces_plot)


calculate_response <- function(data, bl_tw, resp_tw, param) {
    data |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA - 
            mean(current_pA[between(time_s, bl_tw[1], bl_tw[2])])
    ) |>
    filter(between(time_s, resp_tw[1], resp_tw[2])) |>
    summarise(
        parameter = case_when(
            param == "amplitude" ~ min(current_pA),
            param == "area" ~ sum(current_pA) * 0.00005,
            .unmatched = "error"
            )
    )
}

bl_mono1_tw <- c(0.3035, 0.3045)
mono1_tw <- c(0.3075, 0.3085)

AC_mono1_aplitudes <- calculate_response(
    AC_responses,
    bl_tw = bl_mono1_tw,
    resp_tw = mono1_tw,
    param = "amplitude"
    )

bl_mono2_tw <- c(0.3075, 0.3085)
mono2_tw <- c(0.310, 0.314)

AC_mono2_aplitudes <- calculate_response(
    AC_responses,
    bl_tw = bl_mono2_tw,
    resp_tw = mono2_tw,
    param = "amplitude"
    )

bl_mono3_tw <- c(0.316, 0.317)
mono3_tw <- c(0.3175, 0.320)

AC_mono3_aplitudes <- calculate_response(
    AC_responses,
    bl_tw = bl_mono3_tw,
    resp_tw = mono3_tw,
    param = "amplitude"
    )

AC_areas <- calculate_response(
    AC_responses,
    bl_tw = bl_area_tw,
    resp_tw = area_tw,
    param = "area"
    )

AC_areas <- AC_areas |>
    rename(area_pAs = parameter)

AC_mono1_aplitudes <- AC_mono1_aplitudes |>
    rename(mono1_amp_pA = parameter)

AC_mono2_aplitudes <- AC_mono2_aplitudes |>
    rename(mono2_amp_pA = parameter)

AC_mono3_aplitudes <- AC_mono3_aplitudes |>
    rename(mono3_amp_pA = parameter)

AC_all_params <- AC_areas |> 
    left_join(AC_mono1_aplitudes) |>
    left_join(AC_mono2_aplitudes) |>
    left_join(AC_mono3_aplitudes) |>
    left_join(passive_parameters)

write_tsv(AC_all_params, file = "../AC_all_params.tsv")







```



```{r old_calc}
# AC_mono1_amplts <- AC_responses |>
#     group_by(file, sweep) |>
#     mutate(
#         current_pA = current_pA - 
#             mean(current_pA[between(time_s, bl_mono1_tw[1], bl_mono1_tw[2])])
#     ) |>
#     filter(between(time_s, mono1_tw[1], mono1_tw[2])) |>
#     summarise(
#         amplitude_pA = min(current_pA)
#     )
# 
# AC_areas <- AC_responses |>
#     group_by(file, sweep) |>
#     mutate(
#         current_pA = current_pA - 
#             mean(current_pA[between(time_s, bl_area_tw[1], bl_area_tw[2])])
#     ) |>
#     filter(between(time_s, area_tw[1], area_tw[2])) |>
#     summarise(
#         total_area_pAs = sum(current_pA) * 0.00005
#     )

```


### postsynaptic responces to C stimulation

```{r Cpostsynaptic}



```



## Filtering our data from parameters 

```{r filtering}

# selecting only the most important variables
passive_parameters <- passive_parameters |>
      select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF)

baseline_lims <- c(-200,0)
Rs_lims_norm <- c(
    0.9 * median(passive_parameters$Rs_MOhm), 
    1.1 * median(passive_parameters$Rs_MOhm)
    )
Rm_lims <- c(
    0.5 * median(passive_parameters$Rmemb_MOhm), 
    2 * median(passive_parameters$Rmemb_MOhm)
    )
Cm_lims <- c(
    0.5 * median(passive_parameters$Cmemb_pF), 
    2 * median(passive_parameters$Cmemb_pF)
    )

ggplot(passive_parameters, aes(y = baseline_pA))+
    geom_boxplot() + 
    geom_point(aes(group = interaction(file, sweep), x = 0, color = file),
               position = position_dodge(width = 0.15), size = 2.2
               )

```






## Session info

```{r info}

sessionInfo()

```








