---
title: "Import Data 2023-03-07"
author: "Andrii Hovorov"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: false
execute:
  warning: false
  message: false
---

## About

This document imports and processes patch-clamp recording data for a single cell 
(2023_03_07). It covers file import, sweep extraction, 
annotation, passive membrane parameter calculation, quality filtering, and 
post-synaptic response quantification for both AC and C fiber stimulation.

### Setup packages

```{r setup}
library(readABF)
library(tidyverse)
library(DT)
library(plotly)
```

## Cell: 2023_03_07

Tests the import pipeline on a single .abf file. Defines `extract_sweep()` to pull 
one sweep as a tidy dataframe and `abf_to_df()` to loop this over all sweeps in a 
file, producing one combined dataframe per recording.

```{r import_file}
# import one example file
path_abf <- "../data/2023-03-07/2023_03_07_0001.abf"
example_file <- readABF(path_abf)

# make data tidy

head(as.data.frame(example_file, sweep = 1))

# define a function for sweeps extraction
extract_sweep <- function(sweep_n, abf) {
  abf %>%
    as.data.frame(sweep = sweep_n) %>% 
    mutate(sweep = sweep_n)
}

# test on example and extract sweep 1
head(extract_sweep(1, example_file))

# map to extract all sweeps from example file
sweeps_dfs <- map(1:length(example_file$data), ~ extract_sweep(sweep_n = .x, abf = example_file))

# bind rows to create a single dataframe
example_df <- bind_rows(sweeps_dfs)

# define a function to extract all sweeps from different files
abf_to_df <- function(abf) {
  map(1:length(abf$data), extract_sweep, abf = abf) %>%
    bind_rows()
}

# test on example file
head(abf_to_df(example_file))
```

## Now import all the data from the cell

Lists all .abf files in the cell's data folder, imports them, and converts each to a 
tidy dataframe using the functions defined above. Renames columns to `time_s` and 
`current_pA`, drops the voltage channel, and merges all files into a single dataframe 
(`all_sweeps`), tagging each row with its source file.

```{r real_import}

# list the files from the target directory
files <- list.files(
  "../data/2023-03-07", 
  pattern = ".abf",  
  full.names = TRUE
  )

# read all files in the directory
abfs <- map(files, readABF)
# convert files into data frames
dfs <- map(abfs, abf_to_df)

# extract file subname to change naming of dataframes
files |> str_sub(-9, -5)

# name dataframes corresponding to their file
names(dfs) <- str_sub(files, -9, -5)

# make better column names (on example)
invisible(example_df |> rename(
  time_s = `Time [s]`,
  current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
))
datatable(head(example_df, n=15))

# make better column names on all dataframes
dfs <- map(
    dfs, 
    rename,  
    time_s = `Time [s]`,  
    current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
    )
# select only necessary columns (drop voltage)
dfs <- map(dfs, select, time_s, current_pA, sweep)

# merge all dataframes into single one 
# and introduce file name variable
all_sweeps <- bind_rows(dfs, .id = "file")

```

## Add parsed annotation from .txt file

Keeps only the files relevant to the Noradrenaline experiment (excludes A-stim files, 
which weren't recorded for this condition). Adds two annotation columns: `stim` 
(AC vs. C fiber stimulation, based on sweep number) and `cond` (control, 
noradrenaline, or washout), based on which file each sweep came from.
(Here we add the variables for stimulation tyoe and condition (control, noradrenaline, washout))

```{r add_annotation}

all_sweeps <- all_sweeps %>%
    # only the files with AC or C stim, there is no recordings of 
    # A-stim in Noradrenaline so no need to import it
    filter (file %in% c("_0001", "_0002", "_0007", "_0011")) %>%
    mutate(
        # type of fibers stimulated in this file / sweep (C or both A and C)
        stim = case_when(
            file %in% c("_0001", "_0002") ~ if_else(sweep <= 10, "AC", "C"),
            file %in% c("_0007", "_0011") ~ if_else(sweep <= 7, "AC", "C")
        ),
        # condition of recording (control, noradrenaline, washout)
        cond = case_when(
            file %in% c("_0001", "_0002", "_0007") ~ "control",
            file %in% c("_0011") ~ "noradrenalin"
        )
    ) %>%
    mutate(
        stim = as.factor(stim),
        cond = as.factor(cond),
    )

```

```{r check_all_sweeps}
# checking the final dataframe
knitr::kable(head(all_sweeps, 10))
```

## Write data into file

Save the annotated, combined sweep data (`all_sweeps`) to a .tsv file for reuse 
without needing to re-run the import.


```{r}

write_tsv(all_sweeps, file = "../all_sweeps.tsv")

```

## Calculate passive parameters

Calculated the next passive parameters:

- **baseline**: mean current before the step
- **peak**: baseline-subtracted current at the capacitive transient
- **plateau_pA**: baseline-subtracted current at steady state
- **Rs_MOhm**: series (access) resistance, from the peak transient
- **Rtotal_MOhm**: total resistance (series + membrane), from the plateau
- **Rmemb_MOhm**: membrane resistance, obtained by subtracting Rs from Rtotal
- **Cmemb_approx_pF**: a simplified approximation using only Rs (valid when Rmemb \>\> Rs, so the parallel combination is dominated by Rs)
- **Cmemb_pF**: a more accurate estimate using the **parallel combination** of Rs and Rmemb (the actual resistance the capacitor charges through), since Rs and Rmemb form a parallel RC circuit from the capacitor's perspective

```{r passive}

voltage_step_data <- all_sweeps |>
    filter(between(time_s, 0.12, 0.22))

baseline_tw <- c(0.14, 0.16)
peak_tw <- c(0.159, 0.165)
plateau_tw <- c(0.180, 0.198)
decay_tw <- c(0.1605, 0.175)

plot_voltage_step <- ggplot(voltage_step_data, 
       aes(x = time_s, 
           y = current_pA, 
           group = interaction(file, sweep),
           color = file)
       ) +
    geom_vline(xintercept = baseline_tw, color = "lightblue") +
    geom_vline(xintercept = peak_tw, color = "pink") +
    geom_vline(xintercept = plateau_tw, color = "green") +
    geom_vline(xintercept = decay_tw, color = "gold") +
    geom_line(alpha = 0.6)

ggplotly(plot_voltage_step)

passive_parameters <- voltage_step_data |> 
    group_by(file, sweep)  |>
    summarise(
        baseline = mean(current_pA[between(
            time_s, 
            baseline_tw[1], 
            baseline_tw[2])]),
        peak = mean(current_pA[between(
            time_s, 
            peak_tw[1], 
            peak_tw[2])]) - baseline,
        plateau_pA = mean(current_pA[between(
            time_s, 
            plateau_tw[1], 
            plateau_tw[2])]) - baseline,
        Rs_MOhm = -10000/peak,
        Rtotal_MOhm = -10000/ plateau_pA,
        Rmemb_MOhm = Rtotal_MOhm - Rs_MOhm
    )

#### This Part of code will be VERY hard to explain ####
# so we will just copypaste it (it must be robust to errors)
# and i will give a general overview what it is doing:

decay_fits <- voltage_step_data |>
    filter(time_s >= decay_tw[1], time_s <= decay_tw[2]) |>
    nest(.by = c(file, sweep)) |>
    # now we have list-column "data" which will be used for modelling
    mutate(
        models = map(
            data,
            possibly(\(x) nls(x,
                              formula = current_pA ~ SSasymp(time_s, Asym, R0, lrc))
                     )
            ),
        # using SSasymp() as in this tutorial:
        # https://douglas-watson.github.io/post/2018-09_exponential_curve_fitting/
        # but wrapping with possible() to still run if exponent does not fit
        coefs = map(models, coef), #extracting coeficients from model
        lrc = map_dbl(coefs, \(x) pluck(x, "lrc", .default = NA)),
        # pluck() extract value of the lrc from all coefs lists.
        tau = 1/exp(lrc)
        # SSasymp calculate natural Logarithm of Rate Constant (lrc),
        # but we need tau which is 1 / rate constant
        ) |>
    select(file, sweep, tau) # only the variable we needed

 #### END OF HARD PART ####

passive_parameters <- passive_parameters |>
    left_join(decay_fits) |>
    mutate(
        Cmemb_approx_pF = tau / Rs_MOhm * 10^6,
        Cmemb_pF = tau / (1/(1/Rs_MOhm + 1/Rmemb_MOhm)) * 10^6
    )

passive_parameters |>
    select(file, sweep, baseline, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    datatable(rownames = FALSE)

passive_parameters |>
    select(file, sweep, baseline, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    na.omit() |>
    filter(file != "_0011") |>
    GGally::ggpairs(
        aes(color = file),
        columns = 3:6
    )

```

### Filtering

Sets acceptable ranges for baseline, Rs, Rmemb, and Cmemb (based on medians), then 
flags each sweep as pass/fail (`QC`) based on whether it falls within these ranges. 
Visualizes all sweeps as relative percentages of their median to check the filtering 
thresholds, then keeps only sweeps that pass QC for downstream analysis.

```{r filtering}
# setting limits
baseline_lims <- c(-200,0)
Rs_lims <- c(
    0.9 * median(passive_parameters$Rs_MOhm, na.rm = TRUE),
    1.1 * median(passive_parameters$Rs_MOhm, na.rm = TRUE)
    )
Rm_lims <- c(
    0.5 * median(passive_parameters$Rmemb_MOhm, na.rm = TRUE),
    2 * median(passive_parameters$Rmemb_MOhm, na.rm = TRUE)
    )
Cm_lims <- c(
    0.5 * median(passive_parameters$Cmemb_pF, na.rm = TRUE),
    2 * median(passive_parameters$Cmemb_pF, na.rm = TRUE)
    )

# selecting only the most important variables
# then do quality check
passive_parameters <- passive_parameters |>
    ungroup() |>
    select(file, sweep, baseline, Rs_MOhm, Rmemb_MOhm, Cmemb_pF)|>
    mutate(
        QC = between(baseline, baseline_lims[1], baseline_lims[2]) &
            between(Rs_MOhm, Rs_lims[1], Rs_lims[2]) &
            between(Rmemb_MOhm, Rm_lims[1], Rm_lims[2]) &
            between(Cmemb_pF, Cm_lims[1], Cm_lims[2])
    )

# visualisation of quality check
plot_filtering <- ggplot(
    ## first transform dataframe to corrrect format
    data = passive_parameters |>
    mutate(
        bl_pc = baseline / median(baseline, na.rm = TRUE),
        rs_pc = Rs_MOhm / median(Rs_MOhm, na.rm = TRUE),
        rm_pc = Rmemb_MOhm / median(Rmemb_MOhm, na.rm = TRUE),
        cm_pc = Cmemb_pF / median(Cmemb_pF, na.rm = TRUE)
    ) |>
    select(file, sweep, bl_pc, rs_pc, rm_pc, cm_pc, QC) |>
        pivot_longer(
        cols = c(bl_pc, rs_pc, rm_pc, cm_pc),
        names_to = "param",
        values_to = "percentage"
        ),
    aes(x = param, y = percentage)
    ) +
    geom_boxplot(outliers = FALSE) +
    geom_point(
        aes(group = interaction(file, sweep), color = QC),
        position = position_dodge(0.2)
        ) +
    geom_line(
        aes(group = interaction(file, sweep)),
        position = position_dodge(0.2),
        alpha = 0.2
        ) +
    # show our filtering thresholds for each column
    annotate(
        "segment",
        x = c(0.6, 1.6, 1.6, 2.6, 2.6, 3.6, 3.6),
        xend = c(1.4, 2.4, 2.4, 3.4, 3.4, 4.4, 4.4),
        y = c(-200/median(passive_parameters$baseline), 0.5, 2, 0.5, 2, 0.9, 1.1),
        yend = c(-200/median(passive_parameters$baseline), 0.5, 2, 0.5, 2, 0.9, 1.1),
        color = "darkblue"
        ) +
    ylim(0, 4)

ggplotly(plot_filtering)

# filtering raw sweeps data
keep <- passive_parameters %>% filter(QC) %>% select(file, sweep)
sweeps_keep <- all_sweeps |>
    inner_join(keep, by = c("file", "sweep"))
```


### Postsynaptic responces to AC stimulation

Isolates sweeps with AC-fiber stimulation and calculates two response metrics: total 
charge transfer (area under the curve) and the amplitude of the first monosynaptic 
component, each baseline-corrected using a pre-response time window. Includes plots 
to verify the chosen time windows and to compare average traces across conditions. 
Results are merged with passive parameters and saved to file.

```{r}

AC_responses <- sweeps_keep |>
    filter(between(time_s, 0.28, 0.4)) |>
    filter(stim == "AC")

# set timw windows for total area of response
bl_area_tw <- c(0.30660, 0.30965)
area_tw <- c(0.31510, 0.39105)

# quick plotly to check time windows
ACtraces_plot <- ggplot(
    AC_responses,
    aes(
        x = time_s,
        y = current_pA,
        group = interaction(file, sweep),
        color = cond
    )) +
    geom_line(
        alpha = 0.6,
        show.legend = FALSE
        ) +
    geom_vline(xintercept = bl_area_tw, color = "lightblue") +
    geom_vline(xintercept = area_tw, color = "pink") 
    
ggplotly(ACtraces_plot)

AC_area <- AC_responses  |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA - mean(current_pA[
            between(time_s, bl_area_tw[1], bl_area_tw[2])
            ])
    ) |>
    filter(between(time_s, area_tw[1], area_tw[2])) |>
    summarise(
        total_area_pAs = sum(current_pA) * 0.00005
    )

bl_mono1_tw <- c(0.31350, 0.31450)
mono1_tw <- c(0.31920, 0.32020)

AC_mono1_amplts <- AC_responses  |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA - mean(current_pA[
            between(time_s, bl_mono1_tw[1], bl_mono1_tw[2])
            ])
    ) |>
    filter(between(time_s, mono1_tw[1], mono1_tw[2])) |>
    summarise(
        mono1_amplitude_pA = min(current_pA)
    )


calculate_response <- function(data, bl_tw, resp_tw, param) {
    data  |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA - mean(current_pA[
            between(time_s, bl_tw[1], bl_tw[2])
            ])
    ) |>
    filter(between(time_s, resp_tw[1], resp_tw[2])) |>
    summarise(
        parameter = case_when(
            param == "amplitude" ~ min(current_pA), 
            param == "area" ~ sum(current_pA) * 0.00005,
            .unmatched = "error"
        )
    )
}

calculate_response(
    AC_responses, 
    bl_tw = bl_mono1_tw,
    resp_tw = mono1_tw,
    param = "amplitude"
    )


AC_all_params <- AC_area |>
    left_join(AC_mono1_amplts) |>
    left_join(passive_parameters)
    
```

Examine plots with condition discrimination, investigate mean sweeps.

```{r}

AC_area_plot <- AC_responses |>
        group_by(file, sweep) |>
        mutate(
        current_pA = current_pA - mean(current_pA[
            between(time_s, bl_area_tw[1], bl_area_tw[2])
            ]))

AC_mean_plot <- AC_area_plot |>
    group_by(cond, time_s) |>
    summarise(
        current_pA = mean(current_pA)
    )

ACtraces_plot_demeaned <- ggplot(
    AC_area_plot,
    aes(
        x = time_s,
        y = current_pA,
        # group = interaction(file, sweep),
        color = cond
    )) +
    geom_line(
        aes(group = interaction(file, sweep)),
        alpha = 0.1,
        show.legend = FALSE
        ) +
    geom_line(
        data = AC_mean_plot,
        aes(group = cond)
    ) +
    geom_vline(xintercept = bl_area_tw, color = "green") +
    geom_vline(xintercept = area_tw, color = "red") +
    labs(
        x = "Time (s)",
        y = "Current (pA)",
        color = "Condition"
    ) 

ggplotly(ACtraces_plot_demeaned)

```

Write data into separate file

```{r}
write_tsv(AC_all_params, file = "../AC_all_params.tsv")
write_tsv(passive_parameters, file = "../passive_parameters.tsv")
```

### Postsynaptic responces to C stimulation

Same analysis as above (total area and monosynaptic amplitude), but for sweeps with 
C-fiber stimulation, using time windows adjusted for the slower C-fiber response 
kinetics.

```{r}

C_responses <- sweeps_keep |>
    filter(between(time_s, 0.28, 0.4)) |>
    filter(stim == "C")

# set timw windows for total area of response
bl_area_tw_c <- c(0.30660, 0.30965)
area_tw_c <- c(0.32095, 0.38915)

# quick plotly to check time windows
Ctraces_plot <- ggplot(
    C_responses,
    aes(
        x = time_s,
        y = current_pA,
        group = interaction(file, sweep),
        color = file
    )) +
    geom_line(
        alpha = 0.6,
        show.legend = FALSE
        ) +
    geom_vline(xintercept = bl_area_tw_c, color = "lightblue") +
    geom_vline(xintercept = area_tw_c, color = "pink")
    
ggplotly(Ctraces_plot)

C_area <- C_responses  |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA - mean(current_pA[
            between(time_s, bl_area_tw_c[1], bl_area_tw_c[2])
            ])
    ) |>
    filter(between(time_s, area_tw_c[1], area_tw_c[2])) |>
    summarise(
        total_area_pAs = sum(current_pA) * 0.00005
    )

bl_mono1_tw_c <- c(0.31666, 0.31980)
mono1_tw_c <- c(0.32640, 0.33115)

C_mono1_amplts <- C_responses  |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA - mean(current_pA[
            between(time_s, bl_mono1_tw_c[1], bl_mono1_tw_c[2])
            ])
    ) |>
    filter(between(time_s, mono1_tw_c[1], mono1_tw_c[2])) |>
    summarise(
        mono1_amplitude_pA = min(current_pA)
    )

calculate_response(
    C_responses, 
    bl_tw = bl_mono1_tw_c,
    resp_tw = mono1_tw_c,
    param = "amplitude"
    )


C_all_params <- C_area |>
    left_join(C_mono1_amplts) |>
    left_join(passive_parameters)
    
```
```{r}

write_tsv(C_all_params, file = "../C_all_params.tsv")

```

## Session info

```{r session_info}
sessionInfo()
```
