---
title: "Electrophysiological Data Analysis - Rat P12 Lamina X"
author: "Anton Pyatochenko"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: true
---

## Cell ID: 2025_03_24

---

### Setup and Dependencies

```{r setup, message=FALSE, warning=FALSE}
# Load required libraries for ABF file parsing, data manipulation, and interactive visualization
library(readABF)
library(tidyverse)
library(plotly)
library(GGally)
library(DT)

```

---

### Core Utility Functions

```{r functions}
#' Extract an individual sweep from a loaded ABF structure
#' @param sweep_n Integer index of the target sweep
#' @param abf Loaded ABF object containing raw data
#' @return A data frame containing time, current/voltage, and sweep metadata
extract_sweep <- function(sweep_n, abf) {
  abf |>
    as.data.frame(sweep = sweep_n) |>
    mutate(sweep = sweep_n)
}

#' Convert all sweeps within an ABF structure into a unified tidy data frame
#' @param abf Loaded ABF object
#' @return Consolidated data frame of all sweeps
abf_to_df <- function(abf) {
  map(1:length(abf$data), extract_sweep, abf = abf) |>
    bind_rows()
}

#' Generalized function to calculate response amplitudes or integrated areas
#' @param data Cleaned synaptic response data frame
#' @param bl_tw Numeric vector defining the baseline window c(start, end)
#' @param resp_tw Numeric vector defining the response measurement window c(start, end)
#' @param param Character string specifying metric type ("amplitude" or "area")
#' @return A summarized data frame with computed response parameters per sweep
calculate_response <- function(data, bl_tw, resp_tw, param) {
  data |>
    group_by(file, sweep) |>
    mutate(
      current_pA = current_pA - mean(current_pA[between(time_s, bl_tw[1], bl_tw[2])], na.rm = TRUE)
    ) |>
    filter(between(time_s, resp_tw[1], resp_tw[2])) |>
    summarise(
      parameter = case_when(
        param == "amplitude" ~ min(current_pA, na.rm = TRUE),
        param == "area"      ~ sum(current_pA, na.rm = TRUE) * 0.00005,
        .default             = NA_real_
      ),
      .groups = "drop"
    )
}

```

---

### Data Import and Tidy Transformation

```{r import_data}
# Locate and import all binary ABF files from the target directory
files <- list.files("../data/2025_03_24/", pattern = ".abf", full.names = TRUE)
ABFs <- map(files, readABF)

# Assign concise file identifiers based on filename substrings
names(ABFs) <- files %>% str_sub(-9, -5)

# Transform ABF list structures into tidy data frames
DFs <- map(ABFs, abf_to_df)

# Standardize non-standard column names across files
DFs <- map(DFs, rename, 
           time_s = `Time [s]`, 
           current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]")))

# Retain only necessary variables to optimize memory usage
DFs <- map(DFs, select, time_s, current_pA, sweep)

```

---

### Data Consolidation, Filtering, and Metadata Assignment

```{r consolidation}
# Combine individual file data frames into a single master table
All_sweeps <- bind_rows(DFs, .id = "file")

# Filter protocol-specific records and assign experimental conditions and fiber stimulation types
All_sweeps <- All_sweeps %>%
    filter(
        file %in% c("_0008", "_0011", "_0013") | 
            (file == "_0010" & sweep <= 15)
    ) %>%
    mutate(
        stim = case_when(
            file %in% c("_0008", "_0011", "_0013") ~ if_else(sweep <= 10, "AC", "C"),
            file == "_0010" ~ "AC"
        ),
        cond = case_when(
            file %in% c("_0008", "_0010") ~ "control",
            file == "_0011" ~ "noradrenalin",
            file == "_0013" ~ "washout"
        )
    ) %>%
    mutate(
        stim = as.factor(stim),
        cond = as.factor(cond)
    )

```

---

### Voltage Step Filtering, Passive Parameters, and Membrane Modeling ($\tau$)

```{r passive}
# Filter voltage-step data within the test pulse temporal window
voltage_step_data <- All_sweeps |>
    filter(between(time_s, 0.12, 0.22))

# Define time intervals for baseline, peak, and plateau metrics
baseline_tw <- c(0.14, 0.155)
peak_tw     <- c(0.160, 0.165)
plateau_tw  <- c(0.180, 0.198)

# Render interactive overview of voltage-step traces
plot_voltage_step <- ggplot(voltage_step_data, aes(x = time_s, y = current_pA, group = interaction(file, sweep), color = file)) +
    geom_line(alpha = 0.6) +
    geom_vline(xintercept = baseline_tw, color = "lightblue") +
    geom_vline(xintercept = peak_tw, color = "pink") +
    geom_vline(xintercept = plateau_tw, color = "lightgreen") +
    theme_minimal() +
    labs(title = "Voltage Step Traces", x = "Time (s)", y = "Current (pA)")

print(ggplotly(plot_voltage_step))

# Compute passive membrane properties (series resistance, total resistance, membrane resistance)
passive_parameters <- voltage_step_data |>
    group_by(file, sweep) |>
    summarise(
        baseline_pA = mean(current_pA[between(time_s, baseline_tw[1], baseline_tw[2])], na.rm = TRUE),
        peak_pA     = min(current_pA[between(time_s, peak_tw[1], peak_tw[2])], na.rm = TRUE) - baseline_pA,
        plateau_pA  = mean(current_pA[between(time_s, plateau_tw[1], plateau_tw[2])], na.rm = TRUE) - baseline_pA,
        Rs_MOhm     = -10000 / peak_pA,
        Rtotal_MOhm = -10000 / plateau_pA,
        Rmemb_MOhm  = Rtotal_MOhm - Rs_MOhm,
        .groups     = "drop"
    )

# Fit exponential decay curve using non-linear least squares (SSasymp) to extract time constant (tau)
decay_tw <- c(0.1605, 0.175)

decay_fits <- voltage_step_data |>
    filter(between(time_s, decay_tw[1], decay_tw[2])) |>
    nest(.by = c(file, sweep)) |>
    mutate(
        models = map(data, possibly(\(x) nls(x, formula = current_pA ~ SSasymp(time_s, Asym, R0, lrc)))),
        coefs  = map(models, coef),
        lrc    = map_dbl(coefs, \(x) pluck(x, "lrc", .default = NA)),
        tau    = 1 / exp(lrc)
    ) |>
    select(file, sweep, tau)

# Merge decay time constants and calculate membrane capacitance metrics
passive_parameters <- passive_parameters |>
    left_join(decay_fits, by = c("file", "sweep")) |>
    mutate(
        Cmemb_approx_pF = tau / Rs_MOhm * 10^6,
        Cmemb_pF        = tau / (1 / (1 / Rs_MOhm + 1 / Rmemb_MOhm)) * 10^6
    )

# Display interactive table of calculated passive parameters
passive_parameters |>
    select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    DT::datatable(rownames = FALSE)

```

---

### Quality Control (QC) Filtering

```{r qc_filtering}
# --- Quality Control (QC) Filtering (Optimized Limits) ---

baseline_lims <- c(-200, 0)

# Послаблюємо межі для Rs (з ±10% до ±25%), щоб файл з норадреналіном не відсіювався через природні коливання
Rs_lims <- c(
    0.75 * median(passive_parameters$Rs_MOhm, na.rm = TRUE),
    1.25 * median(passive_parameters$Rs_MOhm, na.rm = TRUE)
)

# Розширюємо межі для опору мембрани та ємності
Rm_lims <- c(
    0.4 * median(passive_parameters$Rmemb_MOhm, na.rm = TRUE),
    2.5 * median(passive_parameters$Rmemb_MOhm, na.rm = TRUE)
)

Cm_lims <- c(
    0.4 * median(passive_parameters$Cmemb_pF, na.rm = TRUE),
    2.5 * median(passive_parameters$Cmemb_pF, na.rm = TRUE)
)

# Evaluate records against QC boundary criteria
passive_parameters <- passive_parameters |>
    ungroup() |>
    select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    mutate(
        QC = between(baseline_pA, baseline_lims[1], baseline_lims[2]) &
             between(Rs_MOhm, Rs_lims[1], Rs_lims[2]) &
             between(Rmemb_MOhm, Rm_lims[1], Rm_lims[2]) &
             between(Cmemb_pF, Cm_lims[1], Cm_lims[2])
    )

# Visualize normalized parameter distributions for quality control
plot_filtering <- ggplot(
    data = passive_parameters |>
        mutate(
            bl_pc = baseline_pA / median(baseline_pA, na.rm = TRUE),
            rs_pc = Rs_MOhm / median(Rs_MOhm, na.rm = TRUE),
            rm_pc = Rmemb_MOhm / median(Rmemb_MOhm, na.rm = TRUE),
            cm_pc = Cmemb_pF / median(Cmemb_pF, na.rm = TRUE)
        ) |>
        select(file, sweep, bl_pc, rs_pc, rm_pc, cm_pc, QC) |>
        pivot_longer(cols = c(bl_pc, rs_pc, rm_pc, cm_pc), names_to = "param", values_to = "percentage"),
    aes(x = param, y = percentage)
) +
    geom_boxplot(outliers = FALSE) +
    geom_point(aes(group = interaction(file, sweep), color = QC), position = position_dodge(0.2)) +
    geom_line(aes(group = interaction(file, sweep)), position = position_dodge(0.2), alpha = 0.2) +
    ylim(0, 4) +
    theme_minimal() +
    labs(title = "Quality Control Filtering Diagnostics", x = "Parameter", y = "Normalized Value (% of Median)")

print(ggplotly(plot_filtering))

# Filter raw sweep dataset to retain exclusively QC-verified recordings
keep <- passive_parameters |> filter(QC) |> select(file, sweep)
sweeps_keep <- All_sweeps |>
    inner_join(keep, by = c("file", "sweep"))
```

---

### Synaptic Responses Analysis and Condition Comparison

```{r synaptic_analysis}
# Isolate synaptic records within the target time window and filter strictly for AC fiber stimulation
# synaptic_responses <- sweeps_keep |>
#     filter(between(time_s, 0.28, 0.40)) |>
#     filter(stim == "AC")
synaptic_responses <- sweeps_keep |>
    filter(between(time_s, 0.28, 0.40)) |>
    filter(stim == "C")
# Define optimized time windows for baseline stability and response integration
bl_start <- 0.290
bl_end   <- 0.295
ar_start <- 0.307
ar_end   <- 0.380

bl_m_st  <- 0.3140
bl_m_en  <- 0.3148
m_st     <- 0.3150
m_en     <- 0.3300

# Visualize individual AC synaptic traces overlaid with analysis boundaries
synaptic_traces_plot <- ggplot(synaptic_responses, aes(x = time_s, y = current_pA, group = interaction(file, sweep), color = file)) +
    geom_line(alpha = 0.5, show.legend = FALSE) +
    geom_vline(xintercept = bl_start, color = "lightblue", linetype = "dashed") +
    geom_vline(xintercept = bl_end, color = "lightblue", linetype = "dashed") +
    geom_vline(xintercept = ar_start, color = "gold") +
    geom_vline(xintercept = ar_end, color = "gold") +
    theme_minimal() +
    labs(title = "C Synaptic Responses Overview", x = "Time (s)", y = "Current (pA)")

print(ggplotly(synaptic_traces_plot))

# Normalize AC responses and calculate condition-specific averages for comparative visualization
synaptic_area_plot <- synaptic_responses |>
    group_by(file, sweep) |>
    mutate(current_pA = current_pA - mean(current_pA[between(time_s, bl_start, bl_end)], na.rm = TRUE)) |>
    ungroup()

synaptic_average_plot <- synaptic_area_plot |>
    group_by(cond, time_s) |>
    summarise(current_pA = mean(current_pA, na.rm = TRUE), .groups = "drop")

# Render comparative condition traces plot for C stimulation
condition_comparison_plot <- ggplot(synaptic_area_plot, aes(x = time_s, y = current_pA, color = cond)) +
    geom_line(aes(group = interaction(file, sweep)), alpha = 0.1, show.legend = FALSE) +
    geom_line(data = synaptic_average_plot, aes(group = cond), linewidth = 1.2) +
    geom_vline(xintercept = bl_start, color = "lightblue", linetype = "dashed") +
    geom_vline(xintercept = bl_end, color = "lightblue", linetype = "dashed") +
    geom_vline(xintercept = ar_start, color = "gold") +
    geom_vline(xintercept = ar_end, color = "gold") +
    coord_cartesian(ylim = c(-100, 25)) +
    theme_minimal() +
    labs(title = "C Synaptic Responses Comparison by Condition", x = "Time (s)", y = "Current (pA)", color = "Condition")

print(ggplotly(condition_comparison_plot))
```

---

### Data Consolidation and Export

```{r data_consolidation}
# Compute total integrated area for synaptic responses
synaptic_areas <- synaptic_responses |>
    group_by(file, sweep) |>
    mutate(current_pA = current_pA - mean(current_pA[between(time_s, bl_start, bl_end)], na.rm = TRUE)) |>
    filter(between(time_s, ar_start, ar_end)) |>
    summarise(total_area = sum(current_pA, na.rm = TRUE) * 0.00005, .groups = "drop")

# Extract peak monosynaptic amplitudes
synaptic_amplitude <- calculate_response(
    synaptic_responses,
    bl_tw = c(bl_m_st, bl_m_en),
    resp_tw = c(m_st, m_en),
    param = "amplitude"
) |>
    rename(mono_amplitude = parameter)

# Merge passive properties with active synaptic parameters into a final dataset
final_dataset <- passive_parameters |>
    filter(QC) |>
    left_join(synaptic_areas, by = c("file", "sweep")) |>
    left_join(synaptic_amplitude, by = c("file", "sweep"))

# Display final dataset table
final_dataset |>
    DT::datatable(rownames = FALSE)

# Export processed parameters to TSV file
write_tsv(final_dataset, file = "../Cell_2025_03_24_all_params.tsv")

```

---

### Correlation Analysis and Statistical Trends

```{r correlation_matrix, warning=FALSE, message=FALSE}
# Generate pairwise correlation matrix plot using ggpairs
final_dataset |>
    select(file, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF, total_area, mono_amplitude) |>
    na.omit() |>
    GGally::ggpairs(
        aes(color = file, alpha = 0.6),
        columns = 2:7,
        title = "Cell Parameter Correlation Matrix",
        progress = FALSE
    )

```

```{r scatter_trend}
# Scatter plot with linear regression trendline evaluating amplitude vs. total area relationship
plot_correlation <- ggplot(final_dataset, aes(x = mono_amplitude, y = total_area, color = file)) +
    geom_point(size = 3, alpha = 0.7) +
    geom_smooth(aes(group = 1), method = "lm", se = TRUE, color = "black", linetype = "dashed") +
    theme_minimal() +
    labs(
        title = "Monosynaptic Amplitude vs. Total Area Relationship",
        x = "Monosynaptic Amplitude (pA)",
        y = "Total Area (pA*s)"
    )

print(ggplotly(plot_correlation))

```

```{r numeric_correlation}
# Compute numeric Pearson correlation matrix table
numeric_correlation_matrix <- final_dataset |>
    ungroup() |>
    mutate(mono_amplitude = as.numeric(mono_amplitude)) |>
    select(where(is.numeric), -sweep) |>
    cor(use = "pairwise.complete.obs", method = "pearson")

numeric_correlation_matrix |>
    round(3) |>
    as.data.frame() |>
    DT::datatable(options = list(pageLength = 10))

```

