---
title: "Importing Data"
author: "maryna"
date: "July 20-22, 2026"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: true
execute:
  warning: false
  message: false
---

# Patch clamp data import: cell 2023_03_22_2

This Quarto (`.qmd`) document outlines the workflow for importing, tidying, and analyzing raw electrophysiological patch-clamp recordings from **cell 2023_03_22_2**.

```{r setup}
library(readABF)
library(conflicted)
library(dplyr)
library(DT)
library(tidyverse)
library(plotly)

# Set conflict resolutions to prefer dplyr methods over base/stats
conflict_prefer("rename", "dplyr")
conflict_prefer("filter", "dplyr")
conflicts_prefer(base::as.factor)
```

## Single file processing

We begin by importing a single Axon Binary Format (`.abf`) file to verify sweep extraction logic. Because ABF files contain multiple sweeps, each sweep is coerced into a tidy data frame using `as.data.frame(sweep = ...)` and annotated with its corresponding sweep number.

Here, `sqrt` is passed as a function argument rather than executed immediately (hence no trailing parentheses). The `purrr::map()` function iterates over each sweep index, applying `extract_sweep()` and returning a list of individual data frames.

```{r import_file}
example_file <- readABF("../data/2023-03-22_2/2023_03_22_0032.abf")

# Functional sweep extraction helper
extract_sweep <- function(sweep_n, abf) {
  abf |> 
    as.data.frame(sweep = sweep_n) |> 
    # Append the current sweep index as an explicit column for grouping
    mutate(sweep = sweep_n) 
}

# Iterate over all sweeps in the file and combine into a single data frame
sweep_dfs <- map(1:length(example_file$data), extract_sweep, abf = example_file)
example_df <- bind_rows(sweep_dfs)

# Wrap processing logic into a general helper for batch importing
abf_to_df <- function(abf_given) {
  map(1:length(abf_given$data), extract_sweep, abf = abf_given) |> 
    bind_rows()
}

knitr::kable(head(example_df))
```

## Batch data import and standardization

Next, we batch-import all `.abf` recordings associated with this cell directory.

Setting `full.names = TRUE` in `list.files()` preserves the directory paths necessary for file reading, while `pattern = "\\.abf$"` restricts the search to ABF files.

```{r import}
# Retrieve full file paths for all ABF files in the target directory
files <- list.files("../data/2023-03-22_2/", 
                   full.names = TRUE, 
                   pattern = "\\.abf$")

# Read binary files and parse sweeps into tidy data frames
abfs <- map(files, readABF)
dfs <- map(abfs, abf_to_df)

# Extract file name suffixes (e.g., "_0032") to serve as list identifiers
names(dfs) <- files |> 
  str_sub(-9, -5)

# Standardize column naming across datasets. Backticks are required for 
# legacy column names containing spaces and special characters.
example_df <- example_df |> rename(
  time_s = `Time [s]`,
  current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
)

dfs <- map(
  dfs, 
  rename, 
  time_s = `Time [s]`,
  current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
)
dfs <- map(dfs, select, time_s, current_pA, sweep)

# Combine datasets and annotate experimental variables
all_sweeps <- bind_rows(dfs, .id = "file")

all_sweeps <- all_sweeps |> 
  # Retain only recordings with AC- or C-fiber stimulation.
  # (A-fiber stimulation was not recorded during the Noradrenaline condition.)
  filter(file %in% c("_0032", "_0036", "_0037")) |> 
  mutate(
    # Classify fiber stimulation type based on sweep protocol (Sweeps 1-10: AC; Sweeps 11+: C)
    stim = if_else(sweep <= 10, "AC", "C"),
    # Map file identifiers to experimental treatment conditions
    cond = case_when(
      file == "_0032" ~ "control",
      file == "_0036" ~ "noradrenalin",
      file == "_0037" ~ "washout"
    )
  ) |> 
  mutate(
    stim = as.factor(stim),
    cond = as.factor(cond)
  )

glimpse(all_sweeps)

# Export consolidated dataset to tab-separated values format
write_tsv(all_sweeps, file = 'all_sweeps.tsv')
```

## Passive electrical properties

Extract baseline, peak, and steady-state currents during hyperpolarizing voltage steps to evaluate intrinsic passive membrane properties ($R_s$, $R_{total}$, and $R_m$).

```{r passive}
# Extract the time window containing the hyperpolarizing 
# test pulse (0.12 s – 0.22 s)
voltage_step_data <- all_sweeps |> 
  filter(between(time_s, 0.12, 0.22)) 

# Define analysis time windows (tw) in seconds
baseline_tw <- c(0.14, 0.16)
peak_tw     <- c(0.159, 0.165)
plateau_tw  <- c(0.175, 0.18)

# Visualize test pulse responses with time-window boundaries across conditions
plot_voltage_step <- ggplot(voltage_step_data, aes(
  x = time_s, 
  y = current_pA, 
  group = interaction(file, sweep),
  color = file
)) + 
  geom_vline(xintercept = baseline_tw, color = 'lightblue') +
  geom_vline(xintercept = peak_tw, color = 'lightpink') +
  geom_vline(xintercept = plateau_tw, color = 'darkgreen') +
  geom_line(alpha = 0.3)

# Render interactive plot for visual quality control
ggplotly(plot_voltage_step) 

# Compute baseline-subtracted passive properties and resistance metrics
passive_parameters <- voltage_step_data |> 
  group_by(file, sweep) |> 
  summarise(
    # Baseline current prior to voltage step
    baseline_pA = mean(current_pA[between(time_s, baseline_tw[1], baseline_tw[2])]),
    
    # Peak transient current relative to baseline
    peak_pA = min(current_pA[between(time_s, peak_tw[1], peak_tw[2])]) - baseline_pA,
    
    # Steady-state plateau current
    plateau_pA = mean(current_pA[between(time_s, plateau_tw[1], plateau_tw[2])]),
    
    # Series resistance (Rs), total input resistance (Rtotal), 
    # and isolated membrane resistance (Rm)
    Rs_MOhm = -10000 / peak_pA,
    Rtotal_MOhm = -10000 / plateau_pA,
    Rmembrane_MOhm = Rtotal_MOhm - Rs_MOhm,
    .groups = "drop"
  )

# Define time window for fitting transient decay
decay_tw <- c(0.1605, 0.175)

# Fit single exponential decay to compute charging time constant (tau)
decay_fits <- voltage_step_data |>
  filter(time_s >= decay_tw[1], time_s <= decay_tw[2]) |>
  nest(.by = c(file, sweep)) |>
  mutate(
    # Fit asymptotic exponential model using SSasymp 
    # (formula: current ~ SSasymp(time_s, Asym, R0, lrc))
    # Wrapped in possibly() to handle potential non-convergence gracefully
    models = map(
      data,
      possibly(\(df) nls(current_pA ~ SSasymp(time_s, Asym, R0, lrc), data = df))
    ),
    
    # Extract estimated model parameters
    coefs = map(models, coef),
    
    # Safely pluck natural log of rate constant (lrc)
    lrc = map_dbl(coefs, \(x) pluck(x, "lrc", .default = NA)),
    
    # Calculate time constant (tau = 1 / rate constant = 1 / exp(lrc))
    tau = 1 / exp(lrc)
  ) |>
  select(file, sweep, tau)

# Join decay constants and calculate membrane capacitance (Cm)
passive_parameters <- passive_parameters |> 
  left_join(decay_fits, by = c("file", "sweep")) |> 
  mutate(
    # First-order approximation using series resistance alone
    Cmemb_approx_pF = tau / Rs_MOhm * 10^6,
    
    # Full circuit calculation accounting for parallel parallel Rs and Rm
    Cmemb_pF = tau / (1 / (1 / Rs_MOhm + 1 / Rmembrane_MOhm)) * 10^6
  )

# Display tabular summary of key electrophysiological parameters
passive_parameters |> 
  select(file, sweep, baseline_pA, Rs_MOhm, Rmembrane_MOhm, Cmemb_pF) |> 
  datatable(rownames = FALSE)

# Generate pairwise correlation matrix across extracted passive parameters
passive_parameters |>
  select(file, sweep, baseline_pA, Rs_MOhm, Rmembrane_MOhm, Cmemb_pF) |>
  na.omit() |>
  GGally::ggpairs(
    aes(color = file),
    columns = 3:6
  )
```

## Session info

```{r info}
sessionInfo()
```
