---
title: "Cell: 2025_03_12"
author: "Sofiia Kobryna"
date: 2026.07.20
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: true
execute:
  warning: false
  message: false
---

# About

This is .qmd 

```{r}
#install.packages("GGally")
#install.packages("readABF")
library(readABF)
library(tidyverse)
library(DT)
library(GGally)
library(plotly)
```

# Cell

Lets import our file
```{r import_file}

example_file <-
  readABF("../data/2025_03_12/2025_03_12_0006.abf")

#make data tidy

# as.data.frame(example_file, sweep = 1)

extract_sweep <- function(sweep_n, abf) {
  abf %>% 
    as.data.frame(sweep = sweep_n) %>%
    mutate(sweep = sweep_n)
}

# extract_sweep(sweep_n = 1, abf = example_file)

sweep_dfs <- map(1:length(example_file$data), extract_sweep, abf = example_file)

example_df <- bind_rows(sweep_dfs)

abf_to_df <- function(abf){
  map(1:length(abf$data), extract_sweep, abf = abf) |>
    bind_rows()
}

head(abf_to_df(example_file))



```
# Import all data for the cell

```{r real_import}

files <- list.files("../data/2025_03_12", pattern = ".abf", full.names = TRUE)

abfs <- map(files, readABF)

dfs <- map(abfs, abf_to_df)

files |> str_sub(-9,-5)
names(dfs) <- files |> str_sub(-9,-5)

# example_df |> rename(
#   time_s = `Time [s]`, 
#   current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
# )

dfs <- map(
  dfs, 
  rename, time_s = `Time [s]`, 
  current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
  )

dfs <- map(dfs, select, time_s, current_pA, sweep)
all_sweeps <- bind_rows(dfs, .id = "file")



```
# Annotate

```{r}

# 2025-03-12
All_sweeps <- bind_rows(dfs, .id = "file")

All_sweeps <- All_sweeps |>
    # only files with normal protocol, not "DLF"
    # for "_0008" only recordings before NA addition were included
    filter (file %in% c("_0006", "_0009") | (file == "_0008" & sweep <= 10)) |>
    mutate(
        # type of fibers stimulated in this file / sweep (A, C or both)
        # in this cell only AC simulations were performed
        stim = if_else(sweep <= 10, "AC", "C"),
        # condition of recording (control, noradrenaline, washout)
        cond = case_when(
            file %in% c("_0006", "_0008") ~ "control",
            file == "_0009" ~ "noradrenalin"
        )
    ) |>
    mutate(
        stim = as.factor(stim),
        cond = as.factor(cond),
    )

```

## Example data table 

```{r}
datatable(head(example_df))

knitr::kable(head(example_df))

```
# Session info

```{r}

sessionInfo()

```


```{r}
library(tidyverse)
write_tsv(All_sweeps, file = "all_sweeps.tsv")
all_sweeps <- read_tsv("all_sweeps.tsv")

```

# Calculate passive parametrs

We calculated this passive parameters:
baseline (pA), serial resistance

## Voltage step plot

```{r}

voltage_step_data <- all_sweeps |>
  filter(between(time_s, 0.12, 0.22))

baseline_tw <- c(0.14, 0.16)
peak_tw <- c(0.159, 0.165)
plateau_tw <- c(0.180, 0.198)
decay_tw <- c(0.1605, 0.175)

plot_voltage_step <- ggplot(voltage_step_data,
    aes(x = time_s, y = current_pA, group = interaction(file,sweep), color = file)
    )+
  geom_line(alpha = 0.6) +
  geom_vline(xintercept = baseline_tw, color = "lightblue") +
  geom_vline(xintercept = peak_tw, color = "pink") +
  geom_vline(xintercept = plateau_tw, color = "lightgreen")
  geom_vline(xintercept = decay_tw, color = "gold")

ggplotly(plot_voltage_step)

```

## Passive parameters data table

```{r}

passive_parameters <- voltage_step_data |>
  group_by(file,sweep) |>
  summarise(
    baseline_pA = mean(
      current_pA[between(time_s, baseline_tw[1], baseline_tw[2])]
      ),
    peak_pA = min(
      current_pA[between(time_s, peak_tw[1], peak_tw[2])]
      ) - baseline_pA,
    plateau_pA = mean(
      current_pA[between(time_s, plateau_tw[1], plateau_tw[2])]
      ) - baseline_pA,
    Rs_MOhm = -10000 / peak_pA,
    Rtotal_MOhm = -10000 / plateau_pA,
    Rmemb_MOhm = Rtotal_MOhm - Rs_MOhm
    )

#### This Part of code will be VERY hard to explain ####
# so we will just copypaste it (it must be robust to errors)
# and i will give a general overview what it is doing:

decay_fits <- voltage_step_data |>
  filter(time_s >= decay_tw[1], time_s <= decay_tw[2]) |>
  nest(.by = c(file, sweep)) |>
  # now we have list-column "data" which will be used for modelling
  mutate(
    models = map(
      data,
      possibly(\(x) nls(x,
                        formula = current_pA ~ SSasymp(time_s, Asym, R0, lrc))
               )
      ),
# using SSasymp() as in this tutorial:
# https://douglas-watson.github.io/post/2018-09_exponential_curve_fitting/
# but wrapping with possible() to still run if exponent does not fit
coefs = map(models, coef), #extracting coeficients from model
lrc = map_dbl(coefs, \(x) pluck(x, "lrc", .default = NA)),
# pluck() extract value of the lrc from all coefs lists.
tau = 1/exp(lrc)
# SSasymp calculate natural Logarithm of Rate Constant (lrc),
# but we need tau which is 1 / rate constant
) |>
  select(file, sweep, tau) # only the variable we needed

#### END OF HARD PART ####

passive_parameters <- passive_parameters |>
  left_join(decay_fits) |>
  mutate(
    Cmemb_approx_pF = tau / Rs_MOhm * 10^6,
    Cmemb_pF = tau / (1/(1/Rs_MOhm + 1/Rmemb_MOhm)) * 10^6
    )

passive_parameters |>
  select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
  datatable(rownames = FALSE)

passive_parameters |>
  select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
  na.omit() |>
  GGally::ggpairs(
    aes(color = file),
    columns = 3:6
    )

datatable(passive_parameters)

```
## AC traces plot:

```{r}

AC_responses <- all_sweeps|>
  filter(between(time_s, 0.28, 0.4)) |>
  filter(stim == "AC")
bl_area_tw <- c(0.3062, 0.30985)
area_tw <- c(0.316, 0.382)

ACtraces_plot <- ggplot(AC_responses, aes(
  x = time_s,
  y = current_pA,
  group = interaction(file, sweep),
  color = file
  ))+
  geom_line(
    alpha = 0.5,
    show.legend = FALSE) +
  geom_vline(xintercept = bl_area_tw, color = "lightblue") +
  geom_vline(xintercept = area_tw, color = "gold")


ggplotly(ACtraces_plot)

```

## AC area data table:

```{r}
# 
# AC_area <- AC_responses |>
#   group_by(file, sweep) |>
#   mutate(
#     current_pA = current_pA - 
#       mean(current_pA[between(time_s, bl_area_tw[1], bl_area_tw[2])])
#   ) |>
#   filter(between(time_s, area_tw[1], area_tw[2])) |>
#   summarise(
#     total_area = sum(current_pA) * 0.00005
#   )
# 
# datatable(AC_area)


calculate_response <- function(data, bl_tw, resp_tw, param) {
  data |>
  group_by(file, sweep) |>
  mutate(
    current_pA = current_pA - 
      mean(current_pA[between(time_s, bl_tw[1], bl_tw[2])])
  ) |>
  filter(between(time_s, resp_tw[1], resp_tw[2])) |>
  summarise(
    parametr = case_when(
      param == "amplitude" ~ min(current_pA),
      param == "area" ~ sum(current_pA) * 0.00005,
      .unmatched = "error"
      )
  )
}


AC_area <- calculate_response(
  AC_responses,
  bl_tw = bl_area_tw,
  resp_tw = area_tw,
  param = "area"
)

AC_area <- AC_area |>
  rename(area_pAs = parametr)

datatable(AC_area)

# 
# bl_mono1_tw <- c(0.3185, 0.3162)
# mono1_tw <- c(0.31785, 0.3181)
# 
# AC_mono1_amps <- AC_responses |>
#   group_by(file, sweep) |>
#   mutate(
#     current_pA = current_pA - 
#       mean(current_pA[between(time_s, bl_mono1_tw[1], bl_mono1_tw[2])])
#   ) |>
#   filter(between(time_s, mono1_tw[1], mono1_tw[2])) |>
#   summarise(
#     total_area = min(current_pA)
#   )

```

## AC mono 1 amplitudes data table:

```{r}

# calculate_response <- function(data, bl_tw, resp_tw, param) {
#   data |>
#   group_by(file, sweep) |>
#   mutate(
#     current_pA = current_pA - 
#       mean(current_pA[between(time_s, bl_tw[1], bl_tw[2])])
#   ) |>
#   filter(between(time_s, resp_tw[1], resp_tw[2])) |>
#   summarise(
#     parametr = case_when(
#       param == "amplitude" ~ min(current_pA),
#       param == "area" ~ sum(current_pA) * 0.00005,
#       .unmatched = "error"
#       )
#   )
# }

bl_mono1_tw <- c(0.3156, 0.31605)
mono1_tw <- c(0.31785, 0.3181)

AC_mono1_amplitudes <- calculate_response(
  AC_responses,
  bl_tw = bl_mono1_tw,
  resp_tw = mono1_tw,
  param = "amplitude"
)

AC_mono1_amplitudes <- AC_mono1_amplitudes |>
  rename(mono1_amp_pA = parametr)

datatable(AC_mono1_amplitudes)

```

## AC mono 2 amplitudes data table:

```{r}

bl_mono2_tw <- c(0.31985, 0.32020)
mono2_tw <- c(0.32095, 0.3218)

AC_mono2_amplitudes <- calculate_response(
  AC_responses,
  bl_tw = bl_mono2_tw,
  resp_tw = mono2_tw,
  param = "amplitude"
)

AC_mono2_amplitudes <- AC_mono2_amplitudes |>
  rename(mono2_amp_pA = parametr)

datatable(AC_mono2_amplitudes)

```

## AC mono 3 amplitudes data table:

```{r}

bl_mono3_tw <- c(0.32305, 0.3238)
mono3_tw <- c(0.32485, 0.32585)

AC_mono3_amplitudes <- calculate_response(
  AC_responses,
  bl_tw = bl_mono3_tw,
  resp_tw = mono3_tw,
  param = "amplitude"
)

AC_mono3_amplitudes <- AC_mono3_amplitudes |>
  rename(mono3_amp_pA = parametr)

datatable(AC_mono3_amplitudes)

```

## All

```{r}

# AC_area <- AC_area |>
#   rename(area_pAs = parametr)

# AC_mono1_amplitudes <- AC_mono1_amplitudes |>
#   rename(mono1_amp_pA = parametr)

# AC_mono2_amplitudes <- AC_mono2_amplitudes |>
#   rename(mono2_amp_pA = parametr)

# AC_mono3_amplitudes <- AC_mono3_amplitudes |>
#   rename(mono3_amp_pA = parametr)

AC_all_params <- AC_area |>
  left_join(AC_mono1_amplitudes) |>
  left_join(AC_mono2_amplitudes) |>
  left_join(AC_mono3_amplitudes) |>
  left_join(passive_parameters) 

datatable(AC_all_params)

write_tsv(AC_all_params, file = "../AC_all_params.tsv")

```

```{r}



```