---
title: "Import Data 2023-03-07"
author: "Andrii Hovorov"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: false
execute:
  warning: false
  message: false
---

## About

This is Qmd document for importing Noradrenaline project patch data

### Setup packages

```{r setup}
library(readABF)
library(tidyverse)
library(DT)
library(plotly)
```

## Cell: 2023_03_07

First, let's just import one file

```{r import_file}
# import one example file
path_abf <- "../data/2023-03-07/2023_03_07_0001.abf"
example_file <- readABF(path_abf)

# make data tidy

head(as.data.frame(example_file, sweep = 1))

# define a function for sweeps extraction
extract_sweep <- function(sweep_n, abf) {
  abf %>%
    as.data.frame(sweep = sweep_n) %>% 
    mutate(sweep = sweep_n)
}

# test on example and extract sweep 1
head(extract_sweep(1, example_file))

# map to extract all sweeps from example file
sweeps_dfs <- map(1:length(example_file$data), ~ extract_sweep(sweep_n = .x, abf = example_file))

# bind rows to create a single dataframe
example_df <- bind_rows(sweeps_dfs)

# define a function to extract all sweeps from different files
abf_to_df <- function(abf) {
  map(1:length(abf$data), extract_sweep, abf = abf) %>%
    bind_rows()
}

# test on example file
head(abf_to_df(example_file))
```

## Now import all the data from the cell

```{r real_import}

# list the files from the target directory
files <- list.files(
  "../data/2023-03-07", 
  pattern = ".abf",  
  full.names = TRUE
  )

# read all files in the directory
abfs <- map(files, readABF)
# convert files into data frames
dfs <- map(abfs, abf_to_df)

# extract file subname to change naming of dataframes
files |> str_sub(-9, -5)

# name dataframes corresponding to their file
names(dfs) <- str_sub(files, -9, -5)

# make better column names (on example)
invisible(example_df |> rename(
  time_s = `Time [s]`,
  current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
))
datatable(head(example_df, n=15))

# make better column names on all dataframes
dfs <- map(
    dfs, 
    rename,  
    time_s = `Time [s]`,  
    current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
    )
# select only necessary columns (drop voltage)
dfs <- map(dfs, select, time_s, current_pA, sweep)

# merge all dataframes into single one 
# and introduce file name variable
all_sweeps <- bind_rows(dfs, .id = "file")

```

## Add parsed annotation from .txt file

Here we add the variables for stimulation tyoe and condition (control, noradrenaline, washout)

```{r add_annotation}

all_sweeps <- all_sweeps %>%
    # only the files with AC or C stim, there is no recordings of 
    # A-stim in Noradrenaline so no need to import it
    filter (file %in% c("_0001", "_0002", "_0007", "_0011")) %>%
    mutate(
        # type of fibers stimulated in this file / sweep (C or both A and C)
        stim = case_when(
            file %in% c("_0001", "_0002") ~ if_else(sweep <= 10, "AC", "C"),
            file %in% c("_0007", "_0011") ~ if_else(sweep <= 7, "AC", "C")
        ),
        # condition of recording (control, noradrenaline, washout)
        cond = case_when(
            file %in% c("_0001", "_0002", "_0007") ~ "control",
            file %in% c("_0011") ~ "noradrenalin"
        )
    ) %>%
    mutate(
        stim = as.factor(stim),
        cond = as.factor(cond),
    )

```

```{r check_all_sweeps}
# checking the final dataframe
knitr::kable(head(all_sweeps, 10))
```

## Write data into file

```{r}

write_tsv(all_sweeps, file = "all_sweeps.tsv")

```

## Calculate passive parameters

Calculated the next passive parameters:

- **baseline**: mean current before the step
- **peak**: baseline-subtracted current at the capacitive transient
- **plateau_pA**: baseline-subtracted current at steady state
- **Rs_MOhm**: series (access) resistance, from the peak transient
- **Rtotal_MOhm**: total resistance (series + membrane), from the plateau
- **Rmemb_MOhm**: membrane resistance, obtained by subtracting Rs from Rtotal
- **Cmemb_approx_pF**: a simplified approximation using only Rs (valid when Rmemb \>\> Rs, so the parallel combination is dominated by Rs)
- **Cmemb_pF**: a more accurate estimate using the **parallel combination** of Rs and Rmemb (the actual resistance the capacitor charges through), since Rs and Rmemb form a parallel RC circuit from the capacitor's perspective

```{r passive}

voltage_step_data <- all_sweeps |>
    filter(between(time_s, 0.12, 0.22))

baseline_tw <- c(0.14, 0.16)
peak_tw <- c(0.159, 0.165)
plateau_tw <- c(0.180, 0.198)
decay_tw <- c(0.1605, 0.175)

plot_voltage_step <- ggplot(voltage_step_data, 
       aes(x = time_s, 
           y = current_pA, 
           group = interaction(file, sweep),
           color = file)
       ) +
    geom_vline(xintercept = baseline_tw, color = "lightblue") +
    geom_vline(xintercept = peak_tw, color = "pink") +
    geom_vline(xintercept = plateau_tw, color = "green") +
    geom_vline(xintercept = decay_tw, color = "gold") +
    geom_line(alpha = 0.6)

ggplotly(plot_voltage_step)

passive_parameters <- voltage_step_data |> 
    group_by(file, sweep)  |>
    summarise(
        baseline = mean(current_pA[between(
            time_s, 
            baseline_tw[1], 
            baseline_tw[2])]),
        peak = mean(current_pA[between(
            time_s, 
            peak_tw[1], 
            peak_tw[2])]) - baseline,
        plateau_pA = mean(current_pA[between(
            time_s, 
            plateau_tw[1], 
            plateau_tw[2])]) - baseline,
        Rs_MOhm = -10000/peak,
        Rtotal_MOhm = -10000/ plateau_pA,
        Rmemb_MOhm = Rtotal_MOhm - Rs_MOhm
    )

#### This Part of code will be VERY hard to explain ####
# so we will just copypaste it (it must be robust to errors)
# and i will give a general overview what it is doing:

decay_fits <- voltage_step_data |>
    filter(time_s >= decay_tw[1], time_s <= decay_tw[2]) |>
    nest(.by = c(file, sweep)) |>
    # now we have list-column "data" which will be used for modelling
    mutate(
        models = map(
            data,
            possibly(\(x) nls(x,
                              formula = current_pA ~ SSasymp(time_s, Asym, R0, lrc))
                     )
            ),
        # using SSasymp() as in this tutorial:
        # https://douglas-watson.github.io/post/2018-09_exponential_curve_fitting/
        # but wrapping with possible() to still run if exponent does not fit
        coefs = map(models, coef), #extracting coeficients from model
        lrc = map_dbl(coefs, \(x) pluck(x, "lrc", .default = NA)),
        # pluck() extract value of the lrc from all coefs lists.
        tau = 1/exp(lrc)
        # SSasymp calculate natural Logarithm of Rate Constant (lrc),
        # but we need tau which is 1 / rate constant
        ) |>
    select(file, sweep, tau) # only the variable we needed

 #### END OF HARD PART ####

passive_parameters <- passive_parameters |>
    left_join(decay_fits) |>
    mutate(
        Cmemb_approx_pF = tau / Rs_MOhm * 10^6,
        Cmemb_pF = tau / (1/(1/Rs_MOhm + 1/Rmemb_MOhm)) * 10^6
    )

passive_parameters |>
    select(file, sweep, baseline, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    datatable(rownames = FALSE)

passive_parameters |>
    select(file, sweep, baseline, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    na.omit() |>
    filter(file != "_0011") |>
    GGally::ggpairs(
        aes(color = file),
        columns = 3:6
    )

```

### Postsynaptic responces to AC stimulation

```{r}

AC_responses <- all_sweeps |>
    filter(between(time_s, 0.28, 0.4)) |>
    filter(stim == "AC")

# set timw windows for total area of response
bl_area_tw <- c(0.30660, 0.30965)
area_tw <- c(0.31510, 0.39105)

# quick plotly to check time windows
ACtraces_plot <- ggplot(
    AC_responses,
    aes(
        x = time_s,
        y = current_pA,
        group = interaction(file, sweep),
        color = file
    )) +
    geom_line(
        alpha = 0.6,
        show.legend = FALSE
        ) +
    geom_vline(xintercept = bl_area_tw, color = "lightblue") +
    geom_vline(xintercept = area_tw, color = "pink")
    
ggplotly(ACtraces_plot)

AC_area <- AC_responses  |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA - mean(current_pA[
            between(time_s, bl_area_tw[1], bl_area_tw[2])
            ])
    ) |>
    filter(between(time_s, area_tw[1], area_tw[2])) |>
    summarise(
        total_area_pAs = sum(current_pA) * 0.00005
    )

bl_mono1_tw <- c(0.31350, 0.31450)
mono1_tw <- c(0.31920, 0.32020)

AC_mono1_amplts <- AC_responses  |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA - mean(current_pA[
            between(time_s, bl_mono1_tw[1], bl_mono1_tw[2])
            ])
    ) |>
    filter(between(time_s, mono1_tw[1], mono1_tw[2])) |>
    summarise(
        mono1_amplitude_pA = min(current_pA)
    )


calculate_response <- function(data, bl_tw, resp_tw, param) {
    data  |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA - mean(current_pA[
            between(time_s, bl_tw[1], bl_tw[2])
            ])
    ) |>
    filter(between(time_s, resp_tw[1], resp_tw[2])) |>
    summarise(
        parameter = case_when(
            param == "amplitude" ~ min(current_pA), 
            param == "area" ~ sum(current_pA) * 0.00005,
            .unmatched = "error"
        )
    )
}

calculate_response(
    AC_responses, 
    bl_tw = bl_mono1_tw,
    resp_tw = mono1_tw,
    param = "amplitude"
    )


AC_all_params <- AC_area |>
    left_join(AC_mono1_amplts) |>
    left_join(passive_parameters)
    
```

```{r}
write_tsv(AC_all_params, file = "../AC_all_params.tsv")
```


## Session info

```{r session_info}
sessionInfo()
```
