---
title: "Your Title Here"
author: "Your Name"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: true
---

```{r setup}

library(readABF)
library(tidyverse)
library(plotly)
library(GGally)
```



## My cell: 2025_03_25


## **Sweep Extraction:** Extracts an individual sweep from an ABF object and converts it into a data frame with a sweep index column.
## **ABF Conversion:** Iterates through all data sweeps in an ABF structure and binds them into a single unified data frame.

```{r functions}

extract_sweep <- function(sweep_n, abf){
  abf |>
    as.data.frame(sweep = sweep_n) |>
    mutate(sweep = sweep_n)
  }


abf_to_df <- function(abf){
  map(1:length(abf$data), extract_sweep, abf = abf) |>
    bind_rows()
  }


```

## **File Import & Naming:** Scans the data directory for `.abf` files, reads them, and assigns short identifier names.
## **Tidy Transformation:** Converts ABF structures into tidy data frames, standardizes column names, and selects key variables.


```{r import_data}

files <- list.files("../data/2025_03_25/", pattern = ".abf", full.names = TRUE)

ABFs <- map(files, readABF)

# making short name for each object like "_0004"
names(ABFs) <- files %>% str_sub(-9, -5)

# Transforming abf objects to the tidy dataframes
DFs <- map(ABFs, abf_to_df)

# changing names for variables with non-standard symbols
DFs <- map(DFs, rename, 
           time_s = `Time [s]`, 
           current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
           )

# select only important variables:
DFs <- map(DFs, select, time_s, current_pA, sweep)

```

###Data Consolidation & Filtering: Combines individual data frames into a master table and restricts records based on protocol rules.

###Metadata & Factor Labeling: Assigns fiber stimulation types (stim), recording conditions (cond), and converts them into factors.

```{r _2025_03_25}

# 2025-03-25
all_sweeps <- bind_rows(DFs, .id = "file")

all_sweeps <- all_sweeps %>%
    # only files with normal protocol, not "DLF"
  # file "_0007" is wrong type and should be not included
    # for "_0012" only recordings before NA addition were included
    filter (
        file %in% c("_0011", "_0013", "_0014", "_0015") | 
            (file == "_0012" & sweep <= 10)
        ) %>%
    mutate(
        # type of fibers stimulated in this file / sweep (A, C or both)
        # in this cell only AC simulations were performed
        stim = case_when(
            file %in% c("_0007", "_0013", "_0014", "_0015") ~ if_else(sweep <= 10, "AC", "C"),
            file == "_0011" ~ if_else(sweep <= 12, "AC", "C"),
            file == "_0012" ~ "AC"
        ),
        # condition of recording (control, noradrenaline, washout)
        cond = case_when(
            file %in% c("_0007", "_0011", "_0012") ~ "control",
            file == "_0013" ~ "noradrenalin",
            file %in% c("_0014", "_0015") ~ "washout"
        )
    ) %>%
    mutate(
        stim = as.factor(stim),
        cond = as.factor(cond),
    )

```



###Voltage Step Filtering & Visualization: Subsets data within the 0.12–0.22 s time window, sets specific time intervals for the baseline, peak, and plateau, and renders an interactive graph using ggplot2 and plotly.

###Passive Parameter Calculation: Groups data by file and sweep to compute average baseline current, peak and plateau amplitudes, series resistance (Rs_MOhm), total resistance, and membrane resistance (Rmemb_MOhm).

###Non-Linear Decay Curve Modeling (tau): Filters the decay region (decay_tw), nests data via nest(.by = c(file, sweep)), and applies non-linear least squares (nls) using the SSasymp function to calculate the time constant (tau).

###Membrane Capacitance Calculation (Cmemb): Merges the modeling results and computes both approximate (Cmemb_approx_pF) and precise (Cmemb_pF) membrane capacitance metrics accounting for parallel resistance dynamics.



```{r passive}
voltage_step_data <- all_sweeps |>
    filter(between(time_s, 0.12, 0.22))

baseline_tw <- c(0.14, 0.16)
peak_tw <- c(0.159, 0.165)
plateau_tw <- c(0.180, 0.198)


plot_voltage_step <- ggplot(voltage_step_data,
       aes(x = time_s, y = current_pA, group = interaction(file,sweep), color = file)
       )+
    geom_line(alpha = 0.6) +
    geom_vline(xintercept = baseline_tw, color = "lightblue") +
    geom_vline(xintercept = peak_tw, color = "pink") +
    geom_vline(xintercept = plateau_tw, color = "lightgreen")

ggplotly(plot_voltage_step)

passive_parameters <- voltage_step_data |>
    group_by(file,sweep) |>
    summarise(
        baseline_pA = mean(
            current_pA[between(time_s, baseline_tw[1], baseline_tw[2])]
            ),
        peak_pA = min(
            current_pA[between(time_s, peak_tw[1], peak_tw[2])]
            ) - baseline_pA,
        plateau_pA = mean(
            current_pA[between(time_s, plateau_tw[1], plateau_tw[2])]
            ) - baseline_pA,
        Rs_MOhm = -10000 / peak_pA,
        Rtotal_MOhm = -10000 / plateau_pA,
        Rmemb_MOhm = Rtotal_MOhm - Rs_MOhm
        )



decay_tw <- c(0.1605, 0.175)

#### This Part of code will be VERY hard to explain ####
# so we will just copypaste it (it must be robust to errors)
# and i will give a general overview what it is doing:

decay_fits <- voltage_step_data |>
    filter(time_s >= decay_tw[1], time_s <= decay_tw[2]) |>
    nest(.by = c(file, sweep)) |>
    # now we have list-column "data" which will be used for modelling
    mutate(
        models = map(
            data,
            possibly(\(x) nls(x,
                              formula = current_pA ~ SSasymp(time_s, Asym, R0, lrc))
                     )
            ),
        # using SSasymp() as in this tutorial:
        # douglas-watson.github.io/post/2018-09_exponential_curve_fitting
        # but wrapping with possible() to still run if exponent does not fit
        coefs = map(models, coef), #extracting coeficients from model
        lrc = map_dbl(coefs, \(x) pluck(x, "lrc", .default = NA)),
        # pluck() extract value of the lrc from all coefs lists.
        tau = 1/exp(lrc)
        # SSasymp calculate natural Logarithm of Rate Constant (lrc),
        # but we need tau which is 1 / rate constant
        ) |>
    select(file, sweep, tau) # only the variable we needed

#### END OF HARD PART ####







passive_parameters <- passive_parameters |>
  left_join(decay_fits) |>
  mutate(
    Cmemb_approx_pF = tau / Rs_MOhm * 10^6,
    # tau (s) / R (MOhm) = C (uF), *10^6 to convert to pF
    Cmemb_pF = tau / (1/(1/Rs_MOhm + 1/Rmemb_MOhm)) * 10^6
    # actually tau depends on both Rseries and Rmemb,
    # and from Capacitor point of view they are in parallel
  )




# show calculated passive parameters
passive_parameters |>
  select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
  DT::datatable(rownames = FALSE)

# and nice plot to see them
passive_parameters |>
  select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
  na.omit() |>
  GGally::ggpairs(
    aes(color = file),
    columns = 3:6
  )
```



```{r}

AC_responses <- all_sweeps |>
    filter(between(time_s, 0.28, 0.4)) |>
    filter(stim == "AC")

bl_area_tw <- c(0.305, 0.310)
area_tw <- c(0.314, 0.398)

bl_motol_tw <- c(0.3140, 0.3145)
motol_tw <- c(0.3148, 0.325)

ACTtraces_plot <- ggplot(AC_responses, aes(
    x = time_s,
    y = current_pA,
    group = interaction(file, sweep),
    color = file
))+
    geom_line(
        alpha = 0.5,
        show.legend = FALSE
    ) +
    geom_vline(xintercept = bl_area_tw, color = "lightblue") +
    geom_vline(xintercept = area_tw, color = "gold") +
    geom_vline(xintercept = bl_motol_tw, color = "pink") +
    geom_vline(xintercept = motol_tw, color = "red")

ggplotly(ACTtraces_plot)

AC_responses |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA -
            mean(current_pA[between(time_s, bl_area_tw[1], bl_area_tw[2])])
    ) |>
    filter(between(time_s, area_tw[1], area_tw[2])) |>
    summarise(
        total_area = sum(current_pA) * 0.00005,
        .groups = "drop"
    )

calculate_response <- function(data, bl_tw, resp_tw, param) {
    data |>
        group_by(file, sweep) |>
        mutate(
            current_pA = current_pA -
                mean(current_pA[between(time_s, bl_tw[1], bl_tw[2])])
        ) |>
        filter(between(time_s, resp_tw[1], resp_tw[2])) |>
        summarise(
            parameter = case_when(
                param == "amplitude" ~ min(current_pA),
                param == "area" ~ sum(current_pA) * 0.00005,
                .unmatched = "error"
            ),
            .groups = "drop"
        )
}

Ac_mono_amplitude <- calculate_response(
    AC_responses,
    bl_tw = bl_motol_tw,
    resp_tw = motol_tw,
    param = "amplitude"
)
```
```{r}
# AC_mono1_amplts <- AC_responses |>
#     group_by(file, sweep) |>
#     mutate(
#         current_pA = current_pA -
#             mean(current_pA[between(time_s, bl_mono1_tw[1], bl_mono1_tw[2])])
#     ) |>
#     filter(between(time_s, mono1_tw[1], mono1_tw[2])) |>
#     summarise(
#         amplitude_pA = min(current_pA),
#     )
# 
# AC_areas <- AC_responses |>
#     group_by(file, sweep) |>
#     mutate(
#         current_pA = current_pA -
#             mean(current_pA[between(time_s, bl_area_tw[1], bl_area_tw[2])])
#     ) |>
#     filter(between(time_s, area_tw[1], area_tw[2])) |>
#     summarise(
#         total_area = sum(current_pA) * 0.00005,
#     )
```


```{r data_consolidation}

AC_areas <- AC_responses |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA -
            mean(current_pA[between(time_s, bl_area_tw[1], bl_area_tw[2])])
    ) |>
    filter(between(time_s, area_tw[1], area_tw[2])) |>
    summarise(
        total_area = sum(current_pA) * 0.00005,
        .groups = "drop"
    )


Ac_mono_amplitude <- calculate_response(
    AC_responses,
    bl_tw = bl_motol_tw,
    resp_tw = motol_tw,
    param = "amplitude"
) |>
    rename(mono_amplitude = parameter)


final_dataset <- passive_parameters |>
    left_join(AC_areas, by = c("file", "sweep")) |>
    left_join(Ac_mono_amplitude, by = c("file", "sweep"))


final_dataset |>
    DT::datatable(rownames = FALSE)
```







```{r correlation_matrix, warning=FALSE, message=FALSE}

final_dataset |>
    select(file, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF, total_area, mono_amplitude) |>
    na.omit() |>
    GGally::ggpairs(
        aes(color = file, alpha = 0.6),
        columns = 2:7,
        title = "Матриця співвідношень параметрів клітини",
        progress = FALSE
    )
```







```{r scatter_trend}
plot_correlation <- ggplot(final_dataset, aes(x = mono_amplitude, y = total_area, color = file)) +
    geom_point(size = 3, alpha = 0.7) +
    # Додавання загальної лінії лінійного тренду (без поділу на файли)
    geom_smooth(aes(group = 1), method = "lm", se = TRUE, color = "black", linetype = "dashed") +
    theme_minimal() +
    labs(
        title = "Співвідношення амплітуди та загальної площі",
        x = "Амплітуда моносинаптичної відповіді (pA)",
        y = "Загальна площа (pA*s)"
    )

ggplotly(plot_correlation)
```





```{r numeric_correlation}
numeric_correlation_matrix <- final_dataset |>
    # 1. Зняття групування для коректної роботи select()
    ungroup() |> 
    mutate(
        # 2. Конвертація текстового результату case_when() назад у число
        mono_amplitude = as.numeric(mono_amplitude)
    ) |>
    # Відбір лише числових колонок
    select(where(is.numeric), -sweep) |> 
    # 3. Використання pairwise замість na.omit() для збереження даних
    cor(use = "pairwise.complete.obs", method = "pearson")

# Відображення результату у вигляді таблиці з округленням до 3 знаків
numeric_correlation_matrix |>
    round(3) |>
    as.data.frame() |>
    DT::datatable(options = list(pageLength = 10))
```



