---
title: "Cell ID: 2025_05_02 Report"
author: "Ivan Blashchak"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: true
    
execute:
  warning: false
  message: false
---

## About

This is .Qmd report for patch clamp data analysis
for Lamina X cell (cell ID: 2025_05_02).
Recordings of excitatory postsynaptic currents (EPSCs) 
evoked by primary afferents stimulation were made in control conditions 
and after the addition of Noradrenaline


```{r setup}

# install.packages("readABF")
library(readABF)
library(tidyverse)
library(plotly)
library(DT)

extract_sweep <- function(sweep_n, abf){
    abf |>
        as.data.frame(sweep = sweep_n) |>
        mutate(sweep = sweep_n)
}

abf_to_df <- function(abf){
    map(1:length(abf$data), extract_sweep, abf = abf) |>
        bind_rows()
}

```

## Data import

Files read:

``` {r import}

files <- list.files(
    "../data/2025_05_02/", 
    pattern = ".abf", 
    full.names = TRUE
    )

abfs <- map(files, readABF)

dfs <- map(abfs, abf_to_df)

files |> str_sub(-19,-5)

names(dfs) <- str_sub(files, -9,-5)

dfs <- map(
    dfs, 
    rename, time_s = `Time [s]`, 
    current_pA = any_of(c("Currnt0Pr [pA]", "Volt0Pr [pA]"))
    )

dfs <- map(dfs, select, time_s, current_pA, sweep)

all_sweeps <- bind_rows(dfs, .id = "file")

```

**Part of the imported data.**

After import each file|sweep combination was annotated: 
stimulation type (A, C or both A and C fiber stimulation) and 
condition (control, noradrenaline or washout)

Here is just first 100 datapoints to check:

```{r annotate}


all_sweeps <- all_sweeps %>%
    # for "_0006" only recordings before NA addition were included
    filter ( ! (file == "_0006" & sweep > 10)) %>%
    mutate(
        # type of fibers stimulated in this file / sweep (A, C or both)
        # in this cell only AC simulations were performed
        stim = case_when(
            file == "_0005" ~ if_else(sweep <= 17, "AC", "C"),
            file == "_0006" ~ "C",
            file == "_0007" ~ if_else(sweep <= 18, "AC", "C"),
            file == "_0008" ~ "AC"
        ),
        # condition of recording (control, noradrenaline, washout)
        cond = case_when(
            file %in% c("_0005", "_0006") ~ "control",
            file == "_0007" ~ "noradrenalin",
            file == "_0008" ~ "washout"
        )
    ) %>%
    mutate(
        stim = as.factor(stim),
        cond = as.factor(cond),
    )

datatable(head(all_sweeps, 100))

```

Raw data was saved in `2025_05_02_all_sweeps.tsv`

```{r saving data}

write_tsv(all_sweeps, file = "../2025_05_02_all_sweeps.tsv")
# 
# # saving only part of the data for shiny app to make it faster
# write_tsv(all_sweeps |> filter(between(time_s, 0.1, 0.7)), 
# file = "../all_sweeps_part.tsv")
```


## Passive parameters

From the voltage step (*-10 mV*) at the beginning of each sweep 
we calculated the next parameters of the recording:

- baseline (*pA*)
- serial resistance (*MOhm*)
- membrane resistance (*MOhm*) 
- membrane capacitance (*pF*)

Here are selected time frames to find the baseline (*blue*), peak current (*red*),
plateau current (*green*) and rate constant of the exponential decay (*yellow*)

```{r passive}

# saving only part of data with voltage step to work with
voltage_step_data <- all_sweeps |>
    filter(between(time_s, 0.12, 0.22))

#setting time windows
baseline_tw <- c(0.14, 0.16)
peak_tw <- c(0.159, 0.165)
plateau_tw <- c(0.180, 0.198)
decay_tw <- c(0.1605, 0.175)

# checking time windows
plot_voltage_step <- ggplot(voltage_step_data, 
       aes(x = time_s, y = current_pA, group = interaction(file,sweep), color = file)
       )+
    geom_line(alpha = 0.6) +
    geom_vline(xintercept = baseline_tw, color = "lightblue") +
    geom_vline(xintercept = peak_tw, color = "pink") +
    geom_vline(xintercept = plateau_tw, color = "lightgreen")+
    geom_vline(xintercept = decay_tw, color = "gold")
    

ggplotly(plot_voltage_step)


# calculating resistances
passive_parameters <- voltage_step_data |>
    group_by(file,sweep) |>
    summarise(
        baseline_pA = mean(
            current_pA[between(time_s, baseline_tw[1], baseline_tw[2])]
            ),
        peak_pA = min(
            current_pA[between(time_s, peak_tw[1], peak_tw[2])]
            ) - baseline_pA,
        plateau_pA = mean(
            current_pA[between(time_s, plateau_tw[1], plateau_tw[2])]
            ) - baseline_pA,
        Rs_MOhm = -10000 / peak_pA,
        Rtotal_MOhm = -10000 / plateau_pA,
        Rmemb_MOhm = Rtotal_MOhm - Rs_MOhm
        )


decay_tw <- c(0.1605, 0.175)

#### This Part of code will be VERY hard to explain ####
# so we will just copypaste it (it must be robust to errors)
# and i will give a general overview what it is doing:

decay_fits <- voltage_step_data |> 
    filter(time_s >= decay_tw[1], time_s <= decay_tw[2]) |> 
    nest(.by = c(file, sweep)) |>
    # now we have list-column "data" which will be used for modelling
    mutate(
        models = map(
            data, 
            possibly(\(x) nls(x, 
                              formula = current_pA ~ SSasymp(time_s, Asym, R0, lrc))
                     )
            ),
        # using SSasymp() as in this tutorial:
        # https://douglas-watson.github.io/post/2018-09_exponential_curve_fitting/
        # but wrapping with possible() to still run if exponent does not fit
        coefs = map(models, coef), #extracting coeficients from model
        lrc = map_dbl(coefs, \(x) pluck(x, "lrc", .default = NA)), 
        # pluck() extract value of the lrc from all coefs lists.
        tau = 1/exp(lrc) 
        # SSasymp calculate natural Logarithm of Rate Constant (lrc),
        # but we need tau which is 1 / rate constant
        ) |>
    select(file, sweep, tau) # only the variable we needed

#### END OF HARD PART ####

passive_parameters <- passive_parameters |>
    left_join(decay_fits) |>
    mutate(
        Cmemb_approx_pF = tau / Rs_MOhm * 10^6,
        # tau (s) / R (MOhms) = C (uF), *10^6 to convert to pF
        
        Cmemb_pF = tau / (1/(1/Rs_MOhm + 1/Rmemb_MOhm)) * 10^6
        # actually tau depends on both Rseries and Rmemb, 
        # and from Capacitor point of view they are in parallel 
        )

```

Here are the calculated parameters, and 
their pairs plot to make quick visual check
*NA and outliers are removed from pairs plot*

```{r show_parameters}

# show calculated passive parameters
passive_parameters |>
    select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    datatable(rownames = FALSE)

# and nice plot to see them
passive_parameters |>
      select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF) |>
    na.omit() |>
    filter(baseline_pA >= -500) |>
    GGally::ggpairs(
          aes(color = file), 
          columns = 3:6
          )
```

## Quality filtering 

Next the sweeps was filtered.
To pass quality check the sweep must:

- have *baseline* current within -200 to 0 *pA*
- have *series resistance * within  +- 20% from the median value
- have *membrane resistance*, and *capacitance* within 0.5x to 2x from its median values


```{r filtering}

# setting limits
baseline_lims <- c(-200,0)
Rs_lims <- c(
    0.8 * median(passive_parameters$Rs_MOhm, na.rm = TRUE), 
    1.2 * median(passive_parameters$Rs_MOhm, na.rm = TRUE)
    )
Rm_lims <- c(
    0.5 * median(passive_parameters$Rmemb_MOhm, na.rm = TRUE), 
    2 * median(passive_parameters$Rmemb_MOhm, na.rm = TRUE)
    )
Cm_lims <- c(
    0.5 * median(passive_parameters$Cmemb_pF, na.rm = TRUE), 
    2 * median(passive_parameters$Cmemb_pF, na.rm = TRUE)
    )

# selecting only the most important variables
# then do quality check
passive_parameters <- passive_parameters |>
    ungroup() |>
    select(file, sweep, baseline_pA, Rs_MOhm, Rmemb_MOhm, Cmemb_pF)|>
    mutate(
        QC = between(baseline_pA, baseline_lims[1], baseline_lims[2]) &
            between(Rs_MOhm, Rs_lims[1], Rs_lims[2]) &
            between(Rmemb_MOhm, Rm_lims[1], Rm_lims[2]) &
            between(Cmemb_pF, Cm_lims[1], Cm_lims[2])
    )

# visualisation of quality check
plot_filtering <- ggplot(
    ## first transform dataframe to corrrect format
    data = passive_parameters |>
    mutate(
        bl_pc = baseline_pA / median(baseline_pA, na.rm = TRUE),
        rs_pc = Rs_MOhm / median(Rs_MOhm, na.rm = TRUE),
        rm_pc = Rmemb_MOhm / median(Rmemb_MOhm, na.rm = TRUE),
        cm_pc = Cmemb_pF / median(Cmemb_pF, na.rm = TRUE)
    ) |>
    select(file, sweep, bl_pc, rs_pc, rm_pc, cm_pc, QC) |>
        pivot_longer(
        cols = c(bl_pc, rs_pc, rm_pc, cm_pc), 
        names_to = "param", 
        values_to = "percentage"
        ),
    aes(x = param, y = percentage)
    ) + 
    geom_boxplot(outliers = FALSE) + 
    geom_point(
        aes(group = interaction(file, sweep), color = QC), 
        position = position_dodge(0.2)
        ) + 
    geom_line(
        aes(group = interaction(file, sweep)), 
        position = position_dodge(0.2),
        alpha = 0.05
        ) +
    # show our filtering thresholds for each column
    annotate(
        "segment",
        x = c(0.6, 1.6, 1.6, 2.6, 2.6, 3.6, 3.6),
        xend = c(1.4, 2.4, 2.4, 3.4, 3.4, 4.4, 4.4),
        y = c(-200/median(passive_parameters$baseline_pA), 0.5, 2, 0.5, 2, 0.8, 1.2),
        yend = c(-200/median(passive_parameters$baseline_pA), 0.5, 2, 0.5, 2, 0.8, 1.2),
        color = "darkblue"
        ) +
    scale_x_discrete(labels = c("bl_pc" = "baseline", "rs_pc" = "R series", "rm_pc" = "R membrane", "cm_pc" = "C membrane"))+
    ylim(0, 4) +
    labs(x = "parameter", y = "Normalised values")

ggplotly(plot_filtering)


# filtering raw sweeps data
keep <- passive_parameters %>% filter(QC) %>% select(file, sweep)
sweeps_keep <- all_sweeps |> 
    inner_join(keep, by = c("file", "sweep"))
    

```


## Calculating the postsynaptic responces

### postsynaptic responces to AC stimulation

```{r ACpostsynaptic_setup}

# saving only part of data with voltage step to work with
AC_responses <- sweeps_keep |>
    filter(between(time_s, 0.28, 0.4)) |>
    filter(stim == "AC")

# setting the function to calculate responce
calculate_response <- function(data, bl_tw, resp_tw, param) {
    data |>
    group_by(file, sweep) |>
    mutate(
        current_pA = current_pA - 
            mean(current_pA[between(time_s, bl_tw[1], bl_tw[2])])
    ) |>
    filter(between(time_s, resp_tw[1], resp_tw[2])) |>
    summarise(
        parameter = case_when(
            param == "amplitude" ~ min(current_pA),
            param == "area" ~ sum(current_pA) * 0.00005,
            .unmatched = "error"
            )
    )
}

```

We calculated total area of the response in this cell.
Blue and yellow lines represent time frames 
used to calculate baseline and area.

Solid line shows averaged traces for each conditions.

```{r ACarea}
# setting time windows for total area of response
bl_area_tw <- c(0.285, 0.2935)
area_tw <- c(0.3045, 0.384)

# calculatiing total area
AC_areas <- calculate_response(
    AC_responses,
    bl_tw = bl_area_tw,
    resp_tw = area_tw,
    param = "area"
    ) |>
    rename(area_pAs = parameter)

# making raw and average recordings data for plots
AC_area_plot <- AC_responses |>
        group_by(file,sweep)|>
        mutate(
        current_pA = current_pA - 
            mean(current_pA[between(time_s, bl_area_tw[1], bl_area_tw[2])])
    )

AC_average_plot <- AC_area_plot |>
    group_by(cond, time_s)|>
    summarise(
        current_pA = mean(current_pA)
    )

# plotly to visualise changes in our responce
ACtraces_plot <- ggplot(
    AC_area_plot, 
    aes(
        x = time_s, 
        y = current_pA,
        color = cond
        )
    ) +
    geom_line(
        aes(group = interaction(file, sweep)),
        alpha = 0.1,
        show.legend = FALSE
        ) +
    geom_line(
        data = AC_average_plot,
        aes(group = cond)
        ) +
    geom_vline(xintercept = bl_area_tw, color = "lightblue") +
    geom_vline(xintercept = area_tw, color = "gold") +
    ylim (-100, 20) +
    labs(x = "Time (s)", y = "Current (pA)",
         color = "Condition")

ggplotly(ACtraces_plot)

```

Then we identified three possible monosynaptic responses and calculated their amplitudes.

Blue and yellow lines represent time frames 
used to calculate baseline and amplitude for selected individual mono.

Solid line shows averaged traces for each conditions.


```{R AC monos}

### Mono 1

bl_mono1_tw <- c(0.3035, 0.3045)
mono1_tw <- c(0.3075, 0.3085)

AC_mono1_aplitudes <- calculate_response(
    AC_responses,
    bl_tw = bl_mono1_tw,
    resp_tw = mono1_tw,
    param = "amplitude"
    )

AC_mono1_aplitudes <- AC_mono1_aplitudes |>
    rename(mono1_amp_pA = parameter)

# making raw and average recordings data for plots
AC_mono1_plot <- AC_responses |>
        group_by(file,sweep)|>
        mutate(
        current_pA = current_pA - 
            mean(current_pA[between(time_s, bl_mono1_tw[1], bl_mono1_tw[2])])
    ) |>
    filter(between(time_s, 0.299, 0.315))

AC_average_mono1_plot <- AC_mono1_plot |>
    group_by(cond, time_s)|>
    summarise(
        current_pA = mean(current_pA)
    )

# plotly to visualise changes in our responce
ACtraces_mono1_plot <- ggplot(
    AC_mono1_plot, 
    aes(
        x = time_s, 
        y = current_pA,
        color = cond
        )
    ) +
    geom_line(
        aes(group = interaction(file, sweep)),
        alpha = 0.1,
        show.legend = FALSE
        ) +
    geom_line(
        data = AC_average_mono1_plot,
        aes(group = cond)
        ) +
    geom_vline(xintercept = bl_mono1_tw, color = "lightblue") +
    geom_vline(xintercept = mono1_tw, color = "gold") +
    ylim (-70, 100) +
    xlim(0.299, 0.315) + 
    labs(x = "Time (s)", y = "Current (pA)",
         color = "Condition")

ggplotly(ACtraces_mono1_plot)

### Mono 2

bl_mono2_tw <- c(0.3075, 0.3085)
mono2_tw <- c(0.310, 0.314)

AC_mono2_aplitudes <- calculate_response(
    AC_responses,
    bl_tw = bl_mono2_tw,
    resp_tw = mono2_tw,
    param = "amplitude"
    )

AC_mono2_aplitudes <- AC_mono2_aplitudes |>
    rename(mono2_amp_pA = parameter)

# making raw and average recordings data for plots
AC_mono2_plot <- AC_responses |>
        group_by(file,sweep)|>
        mutate(
        current_pA = current_pA - 
            mean(current_pA[between(time_s, bl_mono2_tw[1], bl_mono2_tw[2])])
    )|>
    filter(between(time_s, 0.299, 0.320))

AC_average_mono2_plot <- AC_mono2_plot |>
    group_by(cond, time_s)|>
    summarise(
        current_pA = mean(current_pA)
    )

# plotly to visualise changes in our responce
ACtraces_mono2_plot <- ggplot(
    AC_mono2_plot, 
    aes(
        x = time_s, 
        y = current_pA,
        color = cond
        )
    ) +
    geom_line(
        aes(group = interaction(file, sweep)),
        alpha = 0.1,
        show.legend = FALSE
        ) +
    geom_line(
        data = AC_average_mono2_plot,
        aes(group = cond)
        ) +
    geom_vline(xintercept = bl_mono2_tw, color = "lightblue") +
    geom_vline(xintercept = mono2_tw, color = "gold") +
    ylim (-70, 100) +
    xlim(0.299, 0.320) + 
    labs(x = "Time (s)", y = "Current (pA)",
         color = "Condition")

ggplotly(ACtraces_mono2_plot)

### Mono 3

bl_mono3_tw <- c(0.316, 0.317)
mono3_tw <- c(0.3175, 0.320)

AC_mono3_aplitudes <- calculate_response(
    AC_responses,
    bl_tw = bl_mono3_tw,
    resp_tw = mono3_tw,
    param = "amplitude"
    )

AC_mono3_aplitudes <- AC_mono3_aplitudes |>
    rename(mono3_amp_pA = parameter)

# making raw and average recordings data for plots
AC_mono3_plot <- AC_responses |>
        group_by(file,sweep)|>
        mutate(
        current_pA = current_pA - 
            mean(current_pA[between(time_s, bl_mono3_tw[1], bl_mono3_tw[2])])
    )|>
    filter(between(time_s, 0.299, 0.325))

AC_average_mono3_plot <- AC_mono3_plot |>
    group_by(cond, time_s)|>
    summarise(
        current_pA = mean(current_pA)
    )

# plotly to visualise changes in our responce
ACtraces_mono3_plot <- ggplot(
    AC_mono3_plot, 
    aes(
        x = time_s, 
        y = current_pA,
        color = cond
        )
    ) +
    geom_line(
        aes(group = interaction(file, sweep)),
        alpha = 0.1,
        show.legend = FALSE
        ) +
    geom_line(
        data = AC_average_mono3_plot,
        aes(group = cond)
        ) +
    geom_vline(xintercept = bl_mono3_tw, color = "lightblue") +
    geom_vline(xintercept = mono3_tw, color = "gold") +
    ylim (-70, 100) +
    xlim(0.299, 0.325) + 
    labs(x = "Time (s)", y = "Current (pA)",
         color = "Condition")

ggplotly(ACtraces_mono3_plot)

```

Areas and amplitudes data was saved in "AC_all_params.tsv"

```{r ACsave}


AC_all_params <- AC_areas |> 
    left_join(AC_mono1_aplitudes) |>
    left_join(AC_mono2_aplitudes) |>
    left_join(AC_mono3_aplitudes) |>
    left_join(passive_parameters)

datatable(AC_all_params)

write_tsv(AC_all_params, file = "../AC_all_params.tsv")


```


### postsynaptic responces to C stimulation

```{r Cpostsynaptic_setup}

# saving only part of data with voltage step to work with
C_responses <- sweeps_keep |>
    filter(between(time_s, 0.28, 0.4)) |>
    filter(stim == "C")

```

We calculated total area of the response in this cell.
Blue and yellow lines represent time frames 
used to calculate baseline and area.

Solid line shows averaged traces for each conditions.

```{r Carea}
# setting time windows for total area of response
bl_area_tw <- c(0.285, 0.2935)
area_tw <- c(0.3045, 0.384)

# calculatiing total area
C_areas <- calculate_response(
    C_responses,
    bl_tw = bl_area_tw,
    resp_tw = area_tw,
    param = "area"
    ) |>
    rename(area_pAs = parameter)

# making raw and average recordings data for plots
C_area_plot <- C_responses |>
        group_by(file,sweep)|>
        mutate(
        current_pA = current_pA - 
            mean(current_pA[between(time_s, bl_area_tw[1], bl_area_tw[2])])
    )

C_average_plot <- C_area_plot |>
    group_by(cond, time_s)|>
    summarise(
        current_pA = mean(current_pA)
    )

# plotly to visualise changes in our responce
Ctraces_plot <- ggplot(
    C_area_plot, 
    aes(
        x = time_s, 
        y = current_pA,
        color = cond
        )
    ) +
    geom_line(
        aes(group = interaction(file, sweep)),
        alpha = 0.1,
        show.legend = FALSE
        ) +
    geom_line(
        data = C_average_plot,
        aes(group = cond)
        ) +
    geom_vline(xintercept = bl_area_tw, color = "lightblue") +
    geom_vline(xintercept = area_tw, color = "gold") +
    ylim (-100, 20) +
    labs(x = "Time (s)", y = "Current (pA)",
         color = "Condition")

ggplotly(Ctraces_plot)

```





## Session info

```{r info}

sessionInfo()

```








